### 2015-11-05 wave_26oct/presentation/methods_results_5_nov.Rmd
Analyze data from pilot study

---
title: "Methods Results 5 November"
author: "Daniel Wall"
date: "November 5, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
setwd("/git_repositories/timeratios/wave_26oct/")
source("load_packages.R")
rm(list = ls())
load("analyzed_data/analyzed_data.Rdata")
```

# Summary

After cleaning the data we had a total of `r n_distinct(pzwhl$responseid)` participants.

## Items that the participants saw:

Note: These are measured on a 0-100 scale.

### Small ratio change

![alt text](small_ratio_change.png)

### Large ratio change

![alt text](big_ratio_change.png)

# Results

## Multilevel model predicting perceived dissimilarity between the items

I ran a linear multilevel model with varying intercepts for subject. Main effects of ratio dissimilarity type (small or large) `rs` and `rb` in the model, item (adding 0, 10, or 20 days to the initial amount) `item1`, `item2`, `item3` in the model, and their interaction. All variables are dummy coded

### Anova with type 3 sums of squares

```{r, echo=FALSE}
Anova(diss_lmer, type = 3)
```

We see a main effect of ratio dissimilarity type, item, and their interaction.

### Model Summary

```{r, echo=FALSE}
summary(diss_lmer)
```

`diss_typers` is the estimate of large ratio dissimilarity minus the estimate of small ratio dissimilarity, while linearly adjusting for all other effects.

`item2` is the estimate of the 0 day change in time minus the 10 day increase in time, while linearly adjusting for all other effects.

### interaction graph for zero type and variance type

The x axis is the perceived dissimilarity between the measures.

```{r, echo=FALSE}

ggplot(eff_df, aes(x = item, y = fit, group = diss_type, color = diss_type)) +
  geom_line() +
  geom_errorbar(aes(ymin = lower, ymax = upper))

```

We see no significant change in percieved dissimilarity by adding time to both delays when ratios are small, but a decrease in percieved dissimilarity by adding time to both delays when ratios are large.

### 2015-11-05 wave_26oct/0_gather_clean_data.R

@@ -1,0 +1,46 @@
+#--------------
+# Gather and clean data for the titrated values for ITC uncertainty
+# Daniel Wall
+# 3 November 2015
+#--------------
+
+setwd("/git_repositories/timeratios/wave_26oct/")
+source("load_packages.R")
+rm(list = ls())
+
+
+#----read in the data----
+
+pzwhl <- read_csv("raw_data/prizewheel_21Oct.csv")
+
+# change column names
+names(pzwhl) <- tolower(names(pzwhl))
+
+# remove test runs
+pzwhl <- pzwhl %>%
+  filter(finished == 1)
+
+# get data so I can work on it for when people are switching
+diss <- pzwhl[, str_detect(names(pzwhl), "(responseid|rs_\\d_\\d|rb_\\d_\\d)")] %>%
+  melt(id = "responseid") %>%
+  mutate(diss_type = str_extract(variable, "(rs|rb)") %>%
+           factor,
+         item = str_extract(variable, "\\d") %>% 
+           factor)
+
+# mixed effects model
+diss_lmer <- lmer(value ~ diss_type * item + (1|responseid),
+                  diss)
+
+Anova(diss_lmer, type = 3)
+summary(diss_lmer)
+
+eff_df <- Effect(c("diss_type", "item"), diss_lmer) %>%
+  as.data.frame()
+
+ggplot(eff_df, aes(x = item, y = fit, group = diss_type, color = diss_type)) +
+  geom_line() +
+  geom_errorbar(aes(ymin = lower, ymax = upper))
+
+
+save.image("analyzed_data/analyzed_data.Rdata")

### 2015-11-05 wave_26oct/load_packages.R

@@ -1,0 +1,16 @@
+#--------------
+# Load packages
+# Daniel Wall
+# 3 November 2015
+#--------------
+library(plyr)
+library(dplyr)
+library(stringr)
+library(readr)
+library(reshape2)
+library(broom)
+library(plotrix)
+library(ggplot2)
+library(lme4)
+library(effects)
+library(car)

### 2015-11-13 shinyapps/scholtenReadSanborn2014.Rmd
Create Shiny app for SRS

---
title: "Scholten Read Sanborn"
author: "Daniel Wall"
date: "November 12, 2015"
output: html_document
runtime: shiny
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(shinyapps)
library(dplyr)
```

```{r, echo=FALSE}
prob_dist <- function (p, alpha) {
  exp(-(-log(p))^alpha)
}

prob_df <- data.frame(prob = seq(0, 1, .01))

```

## Exponential Discounting

In the exponential discounting model, the decision maker will be indifferent between SS and LL when

$$\delta^{t_S}x_S = \delta^{t_L}x_L$$

Where $0 < \delta < 1$ is a one period discount factor and $\rho = 1 - \delta$

One period discount factor, an empiricial measure of discounting over that interval:

$$\delta = (\frac{x_S}{x_L})^{\frac{1}{t_L - t_S}}$$

how does the on period discount factor change for different values  of smaller sooner amount/time, and larger later amount/time.

```{r, echo=FALSE}
exp_disfac <- function(ssamt, sstime, llamt, lltime) {
  (ssamt/llamt)^(1/(lltime - sstime))
}

```

```{r, echo=FALSE}
inputPanel(
  sliderInput("ssamt", "Smaller Sooner $", min=0, max = 500, step=5, value= 10),
    sliderInput("llamt", "Larger Later $", min=0, max = 500, step=5, value= 20),
  sliderInput("sstime", "Smaller Sooner time", min=0, max = 365, step=1, value= 1),
    sliderInput("lltime", "Larger Later time", min=0, max = 365, step=1, value= 2)

  )

renderPrint({
exp_disfac(input$ssamt, input$sstime, input$llamt, input$lltime)

})

```

# Hyperbolic

According to Loewenstein and Prelec’s (1992) hyperbolic discounting model, the deci-
sion maker will be indifferent between SS and LL when

$$d(t_S)v(x_S) = d(t_L)v(x_L)$$

where v is a reference-dependent value function, and d is a generalized hyperbolic dis- count function:

$$d(t) = (1 + \tau t)^{-\beta/\tau}$$

where $\beta > 0$ is delay discounting, and $\tau > 0$ is the departure from exponential discounting. As $\tau$ goes to 0, the generalized hyperbolic discount function reduces to the exponential discount function $d(t)\exp^{-\beta t} = \delta^t$. The hyperbolic discounting model predicts the common difference effect. To see this, the total discounting over the interval tS ? tL can be expressed in the form of a discount **fraction**.

__So discount fraction is different than discount factor__

$$\frac{v(x_S)}{v(x_L)} = (\frac{1+\tau t_S}{1+\tau t_L}) ^{\beta/\tau} = (\frac{1+\tau t_S}{1 + \tau t_S + \tau(t_L-t_x)}) ^{\beta /\tau}$$

```{r}
hyp_disfrac <- function(sstime, lltime, beta, tau) {
  sstime <- as.numeric(sstime)
  lltime <- as.numeric(lltime)
  beta <- as.numeric(beta)
  tau <- as.numeric(tau)

  ((1 + tau * sstime)/(1 + tau * sstime + tau *(lltime - sstime))) ^(beta/tau)
}

hyp_disfrac(0, 1, .1, .1)
```

```{r, echo=FALSE}
inputPanel(
  textInput("sstime", "Smaller Sooner time"),
  textInput("lltime", "Larger Later time"),
  textInput("beta", "beta > 0 delay"),
  textInput("tau", "tau > departure from exponential discounting")
  )

renderPrint({
hyp_disfrac(input$sstime, input$lltime, input$beta, input$tau)

})

```

The value function has 5 properties. First three from Prospect theory

1. Reference Dependence
2. Diminishing Sensitivity
3. Loss aversion

2 additional dealing with elasticity of value function or the ratio between the proportional change in value and the proportional change in amount

$$\epsilon = \frac{v' (x)/v(x)}{1/x}$$

4. Increasing elasticity. For constant proportional increases of x > 0 (decreases of x < 0), v(x) increases (decreases) by increasing proportional amounts. Therefore, if outcome magnitude is increased by a common multiplicative constant, the ratio between the values of xS and xL decreases

5. Loss amplification. Elasticity is greater for losses than for gains. That is, decreasing x < 0 by a given proportion yields a greater propor- tional change in value than increasing x > 0 by the same proportion

```{r}
hyp_val <- function(x, gamma, lambda, mu, sigma) {
  # x <- as.numeric(x)
  if (x >= 0) {
    (1 - gamma) * x^(1-gamma) + (mu *gamma * x) ^ gamma
  } else if (x < 0) {
    -lambda*((1-gamma)*(-x) ^(1-gamma) + (mu + sigma)*gamma*(-x)^gamma)
  } 
}

hyp_val <- function(x, gamma, lambda, mu, sigma) {
  # x <- as.numeric(x)
  ifelse(x >= 0,(1 - gamma) * x^(1-gamma) + (mu *gamma * x) ^ gamma, -lambda*((1-gamma)*(-x) ^(1-gamma) + (mu + sigma)*gamma*(-x)^gamma))

}

hyp_val_df <- data.frame(x = -1000:1000)

hyp_val_df %>% rowwise %>% mutate(value = hyp_val(x, 1, 1, 1, 1))
hyp_val(-10, 1, 1, 1, 1)

```

where ½< c < 1 is diminishing absolute sensitivity to outcomes, l > 0 is increasing elasticity, k > 1 is loss aversion, and r > 0 is loss amplification

```{r, echo=FALSE}
inputPanel(
  sliderInput("gamma", "gamma diminishing absolute sensitivity to outcomes",  min=.1, max = .99, step=.001, value= .1),
  sliderInput("lambda", "Lambda loss aversion",  min=1, max = 4, step=.001, value= .1),
  sliderInput("mu", "mu  increasing elasticity",  min=.1, max = .99, step=.001, value= .1),
  sliderInput("sigma", "sigma loss amplification",  min=.1, max = .99, step=.001, value= .1)
  )

renderPlot({
  hyp_val_df$lambda <- input$lambda
  hyp_val_df$sigma <- input$sigma
  hyp_val_df$mu <- input$mu
  hyp_val_df$gamma <- input$gamma

  hyp_val_df <- transform(hyp_val_df, value = hyp_val(x, gamma, lambda, mu, sigma))

  ggplot(hyp_val_df, aes(x = x, y = value, group = 1)) +
    geom_line() +
    geom_abline(intercept = 0, slope=0) +
    geom_vline(xintercept = 0) +
    xlab("Outcome") +
    ylab("Value") +
    xlim(c(-1000, 1000))
})

```

### 2015-11-13 shinyapps/scholtenReadSanborn2014.Rmd
start on interval model

@@
 }
-
 ```
@@
   sliderInput("lltime", "Larger Later time", min=0, max = 365, step=1, value= 2)
-
   )
@@
 exp_disfac(input$ssamt, input$sstime, input$llamt, input$lltime) 
-
 })
-
 ```
@@
 ```
+
+## The interval discounting model: A hybrid evaluation rule
+
+uses the value function hyperbolic
+
+Interval discount function:
+
+$$D(t_S, t_L) = (1 + \alpha (\frac{w(t_L) - w(t_S)}{\vartheta})^{\vartheta})^{-\beta/\alpha}$$
+
+```{r}
+w_t <- function (t, tau) {
+  t <- as.numeric(t)
+  tau <- as.numeric(tau)
+  (1/tau) * log(1 + tau*t)
+}
+
+int_fn <- function(sstime, lltime, alpha, theta, beta, tau) {
+
+  sstime <- as.numeric(sstime)
+  lltime <- as.numeric(lltime)
+  alpha <- as.numeric(alpha)
+  theta <- as.numeric(theta)
+  beta <- as.numeric(beta)
+
+  (1 + alpha * ((w_t(lltime) - w_t(sstime))/theta)^theta) ^(-beta/alpha)
+}
+
+int_df <- data.frame(int_len = 0:100)
+
+int_df1 <- int_df %>%
+  mutate(sstime =,
+         lltime,
+         alpha,
+         theta,
+         beta)
+
+```
+
+```{r, echo=FALSE}
+inputPanel(
+  textInput("inton", "Interval onset"),
+  textInput("intlen", "Larger Later time"),
+  textInput("beta", "beta > 0 delay"),
+  textInput("tau", "tau > departure from exponential discounting")
+  )
+
+renderPrint({
+hyp_disfrac(input$sstime, input$lltime, input$beta, input$tau)
+
+})
+
+```

### 2015-11-16 shinyapps/scholtenReadSanborn2014.Rmd
Update with interval model and paper introduction

--- 2015-11-13 shinyapps/scholtenReadSanborn2014.Rmd
@@
 ```
+
+## Tradeoff model
+
+attribute-based evaluation rule, in which the outcome advantage of one
+option is weighted against the time advantage of the other, and the option with the great- est advantage is chosen.
+
+These advantages are differences between weighted delays, w(tL) ? w(tS), or the effective interval (as in the interval discounting model), and differences between valued outcomes,
+$v(x_L) - v(x_S)$ for $x_L > x_S > 0$ and $v(x_S) - v(x_L)$ for $x_L < x_S < 0$, or the effective compensation.
+The effective interval and the effective compensation are weighted against one another by a tradeoff function Q.
+
+$$\begin{equation}
+ Q(w(t_L) - w(t_S)) = 
+    \begin{cases}
+      v(x_L) - v(x_S), & \text{if}\ x_L > x_S > 0 \\
+      v(x_S) - v(x_L), & \text{if}\ 0 > x_S > x_L
+    \end{cases}
+  \end{equation}$$
+
+Analogous to the time-weighing function, the trade- off model has a normalized logarithmic value function:
+
+$$\begin{equation}
+ v(x) = 
+    \begin{cases}
+      \frac{1}{\gamma}\log(1+\gamma x), & \text{if}\ x > 0 \\
+      -\frac{\lambda}{\gamma}  \log(1+\gamma (-x)), & \text{if}\ x < 0
+    \end{cases}
+  \end{equation}$$
+
+where $\gamma > 0$ is diminishing absolute sensitivity to outcomes. Like the normalized expo- nential value function discussed by K
+obberling and Wakker (2005), the value function in
+Eq.
+10 approaches constant sensitivity (unit elasticity) as $\gamma$ goes to 0 (linear value func- tion), approaches insensitivity (zero elasticity) as $\gamma$ goes to infinity (zero value function), and is decreasingly elastic otherwise, spanning the entire range from unit elasticity (as x goes to 0) to zero elasticity (as x goes to infinity)
+
+```{r}
+to_val_fn <- function(x, gamma, lambda) {
+  x <- as.numeric(x)
+  gamma <- as.numeric(gamma)
+  lambda <- as.numeric(lambda)
+
+  ifelse(x >= 0, 
+         (1/gamma) * log(1 + gamma * x),
+         -(lambda/gamma) * log(1+gamma*(-x)))
+}
+
+```
+
+```{r}
+inputPanel(
+  sliderInput("gamma_to", "gamma diminishing absolute sensitivity to outcomes",  min=.01, max = .99, step=.001, value= .1),
+  sliderInput("lambda_to", "Lambda loss aversion", value = 1,  min=1, max = 4, step = .05)
+)
+
+renderPlot({
+  to_val_df <- data.frame(x = -100:100)
+  to_val_df$lambda <- input$lambda_to
+  to_val_df$gamma <- input$gamma_to
+
+  to_val_df <- mutate(to_val_df, value = to_val_fn(x, gamma, lambda))
+
+  ggplot(to_val_df, aes(x = x, y = value, group = 1)) +
+    geom_line() +
+    geom_abline(intercept = 0, slope=0) +
+    geom_vline(xintercept = 0) +
+    xlab("Outcome") +
+    ylab("Value") +
+    xlim(c(-100, 100))
+})
+```
+
+The tradeoff fucntion weighs effective intervals against effective compensations
+
+$$Q(t_S, t_L) = \frac{\kappa}{\alpha} \log
+\left(1 + \alpha \left(\frac{w(t_L) - w(t_S)}{\vartheta}\right)^{\vartheta}\right)$$

### 2015-11-16 shinyapps/scholtenReadSanborn2014.Rmd
add interval model

--- 2015-11-13 shinyapps/scholtenReadSanborn2014.Rmd
@@
 
-int_fn <- function(sstime, lltime, alpha, theta, beta, tau) {
+int_fn <- function(int_len, int_onset, alpha, theta, beta, tau) {
 
-  sstime <- as.numeric(sstime)
-  lltime <- as.numeric(lltime)
+  int_len <- as.numeric(int_len)
+  int_onset <- as.numeric(int_onset)
   alpha <- as.numeric(alpha)
@@
 
-  (1 + alpha * ((w_t(lltime) - w_t(sstime))/theta)^theta) ^(-beta/alpha)
+  sstime <- int_onset
+  lltime <- sstime + int_len
+
+  (1 + alpha * ((w_t(lltime, tau) - w_t(sstime, tau))/theta)^theta) ^(-beta/alpha)
 }
 
-int_df <- data.frame(int_len = 0:100)
+int_df <- data.frame(int_len = 1:100)
 
 int_df1 <- int_df %>%
-  mutate(sstime =,
-         lltime,
-         alpha,
-         theta,
-         beta)
+  mutate(int_onset = 1, 
+         alpha = .1, theta = .1, beta = .1, tau = .1)
+
+mutate(int_df1, delta = int_fn(int_len, int_onset, alpha, theta, beta, tau))
 
@@
 inputPanel(
-  textInput("inton", "Interval onset"),
-  textInput("intlen", "Larger Later time"),
-  textInput("beta", "beta > 0 delay"),
-  textInput("tau", "tau > departure from exponential discounting")
+  sliderInput("inton", "Interval onset", min=0, max = 365, step=1, value= 10),
+  sliderInput("alpha", "alpha > 0 subadditivity", min=.01, max = 2, step=.05, value= 0),
+  sliderInput("beta", "beta > 0 delay", min=0, max = 1, step=.05, value= .5),
+  sliderInput("theta", "theta > 1 superadditive", min=1, max = 3, step=.05, value= 1),
+  sliderInput("tau", "tau > 0 diminishing absolute sensitivity", min=.001, max = 1, step=.05, value= .01)
   )
 
-renderPrint({
-hyp_disfrac(input$sstime, input$lltime, input$beta, input$tau)
+renderPlot({
+  int_df$inton <- input$inton
+  int_df$alpha <- input$alpha
+  int_df$beta <- input$beta
+  int_df$theta <- input$theta
+  int_df$tau <- input$tau
 
+  int_df <- transform(int_df, delta = int_fn(int_len, inton, alpha, theta, beta, tau))
+
+    int_df <- transform(int_df, daily = delta ^ (1/int_len))
+
+  ggplot(int_df, aes(x = int_len, y = daily, group = 1)) +
+    geom_line() +
+#     geom_abline(intercept = 0, slope=0) +
+#     geom_vline(xintercept = 0) +
+#     xlab("Outcome") +
+#     ylab("Value") +
+    ylim(c(0, 1))
 })

### 2015-11-17 shinyapps/scholtenReadSanborn2014.Rmd

@@
 
-```{r}
+```{r, echo=FALSE}
 w_t <- function (t, tau) {
@@
          alpha = .1, theta = .1, beta = .1, tau = .1)
-
-mutate(int_df1, delta = int_fn(int_len, int_onset, alpha, theta, beta, tau))
+# 
+# mutate(int_df1, delta = int_fn(int_len, int_onset, alpha, theta, beta, tau))
 
@@
 ```
+
+## Tradeoff model
+
+attribute-based evaluation rule, in which the outcome advantage of one
+option is weighted against the time advantage of the other, and the option with the great- est advantage is chosen.
+
+These advantages are differences between weighted delays, w(tL) ? w(tS), or the effective interval (as in the interval discounting model), and differences between valued outcomes,
+$v(x_L) - v(x_S)$ for $x_L > x_S > 0$ and $v(x_S) - v(x_L)$ for $x_L < x_S < 0$, or the effective compensation.
+The effective interval and the effective compensation are weighted against one another by a tradeoff function Q.
+
+$$\begin{equation}
+ Q(w(t_L) - w(t_S)) = 
+    \begin{cases}
+      v(x_L) - v(x_S), & \text{if}\ x_L > x_S > 0 \\
+      v(x_S) - v(x_L), & \text{if}\ 0 > x_S > x_L
+    \end{cases}
+  \end{equation}$$
+
+Analogous to the time-weighing function, the trade- off model has a normalized logarithmic value function:
+
+$$\begin{equation}
+ v(x) = 
+    \begin{cases}
+      \frac{1}{\gamma}\log(1+\gamma x), & \text{if}\ x > 0 \\
+      -\frac{\lambda}{\gamma}  \log(1+\gamma (-x)), & \text{if}\ x < 0
+    \end{cases}
+  \end{equation}$$
+
+where $\gamma > 0$ is diminishing absolute sensitivity to outcomes. Like the normalized expo- nential value function discussed by K
+obberling and Wakker (2005), the value function in
+Eq.
+10 approaches constant sensitivity (unit elasticity) as $\gamma$ goes to 0 (linear value func- tion), approaches insensitivity (zero elasticity) as $\gamma$ goes to infinity (zero value function), and is decreasingly elastic otherwise, spanning the entire range from unit elasticity (as x goes to 0) to zero elasticity (as x goes to infinity)
+
+```{r,echo=FALSE}
+to_val_fn <- function(x, gamma, lambda) {
+  x <- as.numeric(x)
+  gamma <- as.numeric(gamma)
+  lambda <- as.numeric(lambda)
+
+  ifelse(x >= 0, 
+         (1/gamma) * log(1 + gamma * x),
+         -(lambda/gamma) * log(1+gamma*(-x)))
+}
+
+```
+
+```{r,echo=FALSE}
+inputPanel(
+  sliderInput("gamma_to", "gamma diminishing absolute sensitivity to outcomes",  min=.01, max = .99, step=.001, value= .1),
+  sliderInput("lambda_to", "Lambda loss aversion", value = 1,  min=1, max = 4, step = .05)
+)
+
+renderPlot({
+  to_val_df <- data.frame(x = -100:100)
+  to_val_df$lambda <- input$lambda_to
+  to_val_df$gamma <- input$gamma_to
+
+  to_val_df <- mutate(to_val_df, value = to_val_fn(x, gamma, lambda))
+
+  ggplot(to_val_df, aes(x = x, y = value, group = 1)) +
+    geom_line() +
+    geom_abline(intercept = 0, slope=0) +
+    geom_vline(xintercept = 0) +
+    xlab("Outcome") +
+    ylab("Value") +
+    xlim(c(-100, 100))
+})
+```
+
+The tradeoff fucntion weighs effective intervals against effective compensations
+
+$$Q(t_S, t_L) = \frac{\kappa}{\alpha} \log
+\left(1 + \alpha \left(\frac{w(t_L) - w(t_S)}{\vartheta}\right)^{\vartheta}\right)$$

### 2015-11-18 paper/paperintro.tex
start writing paper for SSRM

\documentclass[]{article}

%opening
\title{An exploration of how ratios affect time}
\author{Daniel Wall}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

When presented with a choice between receiving A) \$10 in 5 days or receiving B) \$12 in 7 days, there are a large number of calculations you could perform to decide which to choose.  Alternative based theories -- such as exponential and hyperbolic discounting -- posit that you calculate a discounted utility for each option. Whereas other models posit that people make attribute based comparisons meaning that they compare the money and time amounts to one another.

\subsection{Exponential Discounting}

The exponential discounting function is  
\begin{equation}\ref{eq:exp}
	\delta_{exp} = \exp^{-rt}
\end{equation}

Where $\delta$ is the discount fraction which each monetary amount is multiplied by and $r$ is the exponential discount rate. And the utlility of an option is:

\begin{equation}\ref{eq:utilalt}
	U(x, t) = x \delta
\end{equation}

 If the daily exponential discount rate is .01 then the current value of A) Would be $\$10 * \exp^{-.01*5} = 9.5$ and the current value of B) would be $\$12 * \exp^{-.01*7} = 11.19$

From \cite{@Toubia2014} we can get the probability of choosing larger later from the following formula

\begin{equation}\label{eq:pll}
	P_{LL} = \frac{\exp(U_{LL})}{\exp(U_{LL})+ \exp(U_{SS})}
\end{equation}

Using this formula we get the following probability of choosing larger later $\frac{11.19}{11.19+9.5} = .54$

One of the empirical regularities in intertemporal choice research is that impatience decreases when times are moved away from the present. The exponential model, however, does not account for this phenomenon. For instance if we added 100 days to both A and B yeilding C) \$10 in 105 days and D) \$ 12 in 107 days the utility of C would be $3.5$ and the utility of D would be $4.11$. Plugging into equation \ref{eq:pll} we get $.54$, the same answer we saw before.

\subsection{Hyperbolic}

Hyperbolic discounting has been shown to account for decreasing impatience

\begin{equation}\label{eq:hyp}
\delta_{hyp} = \frac{1}{1+kt}
\end{equation}

Where $k$ is the hyperbolic discount rate, and the utility of the option is found by equation \ref{eq:utilatt}.

Assuming daily $k = .01$, for choice A) \ref{eq:hyp} we get $\delta_A = \frac{1}{1+.01*5} = .952$ meaning the utility is $9.52$ and for B) the utility is $11.21$ which yields a larger later probability of $.541$. When we look at the larger later probabilities for C and D we get, 4.88 and  5.79 yielding a larger later value of $.543$. There's an increase in the likelihood of choosing the larger later option when both options are moved into the future, this is known as diminishing impatience.

The hyperbolic and exponential models both predict that people make alternative based transitions; however recent models have shown that people make attribute based transitions

\cite{@Loewenstein1991}

\subsection{ITCH}

Marzilli Ericson and colleagues showed that fully attribute based models can predict out of sample choices better than alternative based models. 
According to ITCH the probability of choosing larger later is

\begin{equation}\label{eq:itch}
	P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
\end{equation}

In the paper they give the example parameters for each of the coefficients of 0, .1, .1. -.1, -.1. Using these coefficients in \ref{eq:itch} for A compared to B the probability of choosing larger later would be $.496$, whereas for C and D the probability of choosing larger later would be $.504$

\end{document}

### 2015-11-18 presentation/ITCH ratio.Rmd
start writing paper for SSRM

---
title: "ITCH Predictions"
author: "Daniel Wall"
date: "November 17, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

From @MarzilliEricson2015

According to ITCH the probability of choosing larger later is

$$P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)$$

Where $x^* = (x_1 + x_2)/2$

The function below calculates the probability that a participant will choose larger later.  
```{r}
p_ll <- function(beta_intercept, beta_xa, beta_xr, beta_ta, beta_tr, x1, x2, t1, t2) {
  x_rat <- (x2 - x1)/ ((x2 + x1)/2)
  t_rat <- (t2 - t1)/ ((t2 + t1)/2)
  reg <- beta_intercept + beta_xa*(x2 - x1) + beta_xr*(x_rat) + 
    beta_ta*(t2 - t1) + beta_tr*(t_rat)
  plogis(reg)
} 
```

To confirm that it is right let's use the example from the paper

```
Consider the following concrete example. If the par- ticipant uses weights (0.0, 0.1, 0.1, −0.1, −0.1) and selects between $10 today and $20 tomorrow, then the probabil- ity of choosing the later option is .68
```

```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 20, 0, 1)
```

The function appears to be working

Let's see what is says for our example of $10 in 5 days and $11 in 6 days

```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 6)

```

Now if we add 5 days to both

```{r}

p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 11)

```

$P_{LL}^{10, 11}$   from the $P_{LL}^{5, 6}$ `r p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 11) - p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 6)`.

Now if we use $10 in 5 days and $11 in 15 days'

```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 15)

```

Now by adding 5 days to each

```{r}
p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 20)

```

When subtracting the $P_{LL}^{10, 20}$   from the $P_{LL}^{5, 15}$ we get `r p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 10, 20) - p_ll(0.0, 0.1, 0.1, −0.1, −0.1, 10, 11, 5, 15)`

So ITCH would predict that the there would be a small  difference in decreasing impatience between the two scenarios. I think we weren't accounting for the difference earlier today.

### 2015-11-18 shinyapps/scholtenReadSanborn2014.Rmd
start writing paper for SSRM

@@
 date: "November 12, 2015"
-output: html_document
+output: 
+  html_document: 
+    number_sections: yes
+    toc: yes
 runtime: shiny
@@
 
-## Exponential Discounting
+# Exponential Discounting
 
@@
 
-## The interval discounting model: A hybrid evaluation rule
+# The interval discounting model: A hybrid evaluation rule
 
@@
 
-## Tradeoff model
+# Tradeoff model
 

### 2015-11-18 shinyapps/probability_ll_ITCH.R

@@ -1,0 +1,40 @@
+df <- data.frame(t1 = seq(5, 100, by = 10))
+
+
+library(dplyr)
+
+p_ll <- function(beta_intercept, beta_xa, beta_xr, beta_ta, beta_tr, x1, x2, t1, t2) {
+  x_rat <- (x2 - x1)/ ((x2 + x1)/2)
+  t_rat <- (t2 - t1)/ ((t2 + t1)/2)
+  reg <- beta_intercept + beta_xa*(x2 - x1) + beta_xr*(x_rat) + 
+    beta_ta*(t2 - t1) + beta_tr*(t_rat)
+  plogis(reg)
+} 
+
+p_ll(0, .1, .1, -.1, -.1, 10, 20, 10, 11)
+
+df <- df %>% 
+  mutate(t2 = t1 + 1,
+         x1 = 10, 
+         x2 = 20,
+         beta_intercept = 0, beta_xa = .1, beta_xr  = .1, 
+         beta_ta  = -.1, beta_tr = -.1,
+         p_ll = p_ll(beta_intercept, beta_xa, beta_xr, beta_ta, beta_tr, x1, x2, t1, t2))
+
+df2 <-  data.frame(t1 = seq(1, 100, by = 3))
+
+
+df2 <- df2 %>% 
+  mutate(t2 = t1 + 5,
+         x1 = 10, 
+         x2 = 20,
+         beta_intercept = 0, beta_xa = .1, beta_xr  = .1, 
+         beta_ta  = -.1, beta_tr = -.1,
+         p_ll = p_ll(beta_intercept, beta_xa, beta_xr, beta_ta, beta_tr, x1, x2, t1, t2))
+
+p_ll(0, .1, .1, -.1, -.1, 10, 20, 5, 6)/p_ll(0, .1, .1, -.1, -.1, 10, 20, 10, 11)
+p_ll(0, .1, .1, -.1, -.1, 10, 20, 5, 6)-p_ll(0, .1, .1, -.1, -.1, 10, 20, 10, 11)
+
+p_ll(0, .1, .1, -.1, -.1, 10, 20, 5, 15)/p_ll(0, .1, .1, -.1, -.1, 10, 20, 10, 20)
+p_ll(0, .1, .1, -.1, -.1, 10, 20, 5, 15)-p_ll(0, .1, .1, -.1, -.1, 10, 20, 10, 20)
+

### 2015-11-30 paper/paperintro.tex
Update to paper

@@
 %opening
-\title{An exploration of how ratios affect time}
+\title{An exploration of how ratios affect time discounting}
 \author{Daniel Wall}
@@
 \begin{abstract}
-
+	People use ratios in everyday life. 
+	Current models of intertemporal choice posit that people calculate ratios of both money and time. 
+	These models account for diminishing impatience -- when items are moved into the future, people are more patient. 
+	Specifically we manipulate the size of ratios between times and show that although some models predict that the degree of diminishing impatience will be larger when the difference between the ratio of times is large, we show that these models underestimate the size of this effect.
+	By modifying a model of ITC we show a better fit for our data. 
+	This better fit leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
 \end{abstract}
@@
 \section{Introduction}
+
+A large portion of our daily decisions involve making decisions involving time in some respect, often taking the form of a the tradeoff between receiving something of smaller value now and receiving something of larger value in the future.
+For instance when a consumer selects a shipping speed when shopping online they are making an intertemporal choice.
+These decisions require consumers to, perhaps implicitly, calculate how much waiting for the slower shipping speed is worth.
+
+ Firms could have differing preferences on whether a consumer is impatient or patient when choosing a shipping speed. 
+ If a firm’s inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster.
+ It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
+This understanding how to optimally display information leads to novel predictions for how to display shipping speeds.
+
+Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
+Across ___ studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
+Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios. 
+Depending on the ratio we can make choices more or less patient
+
+\section{Ratios and Intertemporal Choices}
+
+Ratios are ubiquitous in the modern world and are paramount in disparate fields from the stock market to baking. 
+When making inferences from ratios, people are biased in predictable ways.
+For instance, Yamagishi showed that ratios with large numerators and denominators are perceived to be riskier than ratios with small numerators and denominators.
+Further, Reyna and Brainerd argued that people overweight numerators compared to denominators.
+
+Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
+An example intertemporal choice is choosing between \$5 in 1 day and \$6 in 5 days. 
+The ITCH model would predict that participants calculate differences between and ratios of both the dollar amounts and the times. 
+If people are biased in calculating ratios for everyday items, it follows that they would be biased in calculating ratios in intertemporal choices.
+
+One of the empirical regularities in intertemporal choice research is that impatience decreases when times are moved away from the present.
+If we took the above example and moved both options back by 20 days we would see more patient responses. 
+Multiple explanations for this phenomenon -- known as decreasing impatience -- have been marshaled.
+
+Zauberman showed that taking into account subjective perceptions of time can account for decreasing impatience. 
+For instance they find that the subjective time perception of 1 year is not different from the subjective time perception of 3 years. 
+Further subjective time perception was shown to follow a psychophysical logarithmic function.
+Zauberman, however, embedded this subjective time perception into a hyperbolic discounting function – an alternative based discounting function.
+
+Beyond accounting for diminishing impatience subjective time perception can account for subadditive discounting. 
+Read (2001) found that intertemporal choices are intransitive, specifically “discounting over a delay is greater when the delay is divided into subintervals than when it is left undivided”. 
+However, Zauberman showed that participants perceive the total time horizon as longer when it is divided into subintervals than when it is not divided. 
+Further participants discounted more when the delay was divided into subintervals than when it was undivided.
+
+The focus of this paper is on how people calculate ratios specifically about time. 
+We hypothesize that the ratio change in times alters the extent of decreasing impatience. 
+We begin with a review of the current discounting models and how they account for decreasing impatience. 
+Further we will show that while some models predict that ratio change will have an effect on decreasing impatience, that none of the models predict effects as large as we see. 
+These results introduce a phenomenon whose extent is not captured by the current discounting models. 
+They show that the properties of the numbers given can alter patience via their ratio. 
+Moreover recent models of intertemporal choice do not properly predict the degree of decreasing impatience properly, we show that our modification to their model does a better job of predicting choices.
+
+\section{Discounting Models}
 
@@
 The exponential discounting function is
+
 \begin{equation}\ref{eq:exp}
@@
 
-From \cite{@Toubia2014} we can get the probability of choosing larger later from the following formula
+From \cite{@Toubia2014} Toubia 2014 we can get the probability of choosing larger later from the following formula
 
@@
 
-Where $k$ is the hyperbolic discount rate, and the utility of the option is found by equation \ref{eq:utilatt}.
+Where $k$ is the hyperbolic discount rate, and the utility of the option is found by equation \ref{eq:utilalt}.
 
@@
 
-\cite{@Loewenstein1991}
+\subsubsection{Subjective Time Perception}
+
+Zauberman showed that peoples subjective perception of time can account for hyperbolic discounting.
 
@@
 -.1, -.1.
-Using these coefficients in \ref{eq:itch} for A compared to B the probability of choosing larger later would be $.496$, whereas for C and D the probability of choosing larger later would be $.504$
+Using these coefficients in equation \ref{eq:itch} for A compared to B the probability of choosing larger later would be $.496$, whereas for C and D the probability of choosing larger later would be $.504$
+
+\subsection{Discounting By Intervals}
+
+\begin{equation}\label{eq:int}
+	D(t_S, t_L) = \left(1 + \alpha \left(\frac{w(t_L) - w(t_S)}{\vartheta}\right)^{\vartheta}\right)^{-\beta/\alpha}
+\end{equation}
+
+\subsection{Trade-off model}
+
+% %\begin{equation}
+% %Q(w(t_L) - w(t_S)) = 
+% %\begin{cases}
+% %v(x_L) - v(x_S), & \text{if}\ x_L > x_S > 0 \\
+% %v(x_S) - v(x_L), & \text{if}\ 0 > x_S > x_L
+% %\end{cases}
+% %\end{equation}
+
+\subsection{Summary}
+Taken together these models show that some models predict that a large ratio change will have more decreasing impatience than a small ratio change; however we posit that the degree of  decreasing impatience will be larger when the ratio change is small compared to when it is large.
+Take the following two choices for example $SS_{sr}$ = \$10 in 5 days and $LL_{sr}$ = \$12 in 6 days and $SS_{lr}$ = \$10 in 5 days and $LL_{lr}$ = \$12 in 10 days.
+Where $SS_{sr}$ stands for the smaller sooner option for the small time ratio.
+If we add 5 days to both items they would now be $SS_{sr}^{del}$ = \$10 in 10 days and $LL_{sr}^{del}$ = \$12 in 11 days and $SS_{lr}^{del}$ = \$10 in 10 days and $LL_{lr}^{del}$ = \$12 in 15 days.
+
+[Insert a table about the degree of decreasing impatience for each model here]
+
+\section{Time Ratio Model (TRaM)}
+
+We propose a new model of ITC which can account for the degree of decreasing impatience based on the ratios of time. 
+We begin with a verbal description of our model
+
+\subsection{An example of TRM}
+
+When presented with a choice between \$5 in 1 day and \$10 in 5 days. 
+A TRM decision maker would look at the two delays and calcluate their ratio. 
+Then they will use a rule of thumb to say how much relative compensation they need to account for that ratio. Below is an example of a TRM decision maker
+The time is 5 times as long therefore I will need 3 times as much money.
+
+If people use this heuristic then the ratio between times becomes important. 
+If we move both options up by five days then the ratio between times becomes $10/6$ = 1.66. A participant using the TRM heuristic would say that 
+The larger later time is about 1.5 times as long as the smaller sooner time therefore I need 1.25 times as much money.
+
+\section{Study 1}
+
+In study 1 we will test this prediction of the degree of decreasing impatience depending on the change in ratio.
+
+\subsection{Methods}
+Participants were either given 8 choices with either a small or large time ratio which was moved into the future by 5 days.
+
+In condition 1 participants were given 8 intertemporal choices however the focal questions were 
+\$10 in 5 days or \$11 in 6 days
+and 
+\$10 in 10 days or \$11 in 11 days
+
+In condition 1 participants were given 8 intertemporal choices however the focal questions were 
+\$10 in 5 days or \$11 in 6 days
+and 
+\$10 in 10 days or \$11 in 11 days
+
+In condition 2 participants were given 8 intertemporal choices and the focal questions were 
+\$10 in 5 days or \$11 in 15 days
+and 
+\$10 in 10 days or \$11 in 20 days
+
+\subsubsection{Participants}
+
+\subsubsection{Results}
+
+We calculated the daily discount fraction for each of the four focal items while controlling for the participants discount fraction from the other 6 items. 
+The degree of decreasing impatience was calculated by taking the difference between the daily discount fraction from the two focal items. 
+We ran a multilevel linear regression predicting the degree of decreasing impatience with condition and discount fraction of the other items as predictors.
+
+The regression showed that the degree of decreasing impatience was predicted by condition.
 

### 2015-11-30 paper/paperintro.tex
updated version of paper with 4 studies

@@
 \begin{abstract}
-	People use ratios in everyday life. 
-	Current models of intertemporal choice posit that people calculate ratios of both money and time. 
+	Choices involving time, intertemporal choices, are ubiquitous -- for instance choosing a shipping speed for your latest Amazon order. 
+	Further, people use ratios in everyday life -- intertemporal choices are no different, 
+	recent cognitive models of intertemporal choice show that people calculate ratios of both money and time. 
 	These models account for diminishing impatience -- when items are moved into the future, people are more patient. 
-	Specifically we manipulate the size of ratios between times and show that although some models predict that the degree of diminishing impatience will be larger when the difference between the ratio of times is large, we show that these models underestimate the size of this effect.
-	By modifying a model of ITC we show a better fit for our data. 
-	This better fit leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
+	Our goal is twofold: 1) to show that previous models of intertemporal choice do not properly account for temporal ratios and present a conceptual model of time discounting, and  2) by explicitly manipulating how people use ratios of time, we can alter how people make intertemporal choices.
+	This leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
+	Further we show that in a hypothetical shipping scenario that when temporal ratios are explicit, people are more impatient.
+
 \end{abstract}
@@
 
-A large portion of our daily decisions involve making decisions involving time in some respect, often taking the form of a the tradeoff between receiving something of smaller value now and receiving something of larger value in the future.
+Many of our daily choices involve making decisions involving time in some respect, often taking the form of a the tradeoff between receiving something of smaller value now and receiving something of larger value in the future.
 For instance when a consumer selects a shipping speed when shopping online they are making an intertemporal choice.
@@
 Firms could have differing preferences on whether a consumer is impatient or patient when choosing a shipping speed. 
- If a firm’s inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster.
+If a firm's inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster.
 It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
@@
 Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
-Across ___ studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
+Across [] studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
 Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios. 
-Depending on the ratio we can make choices more or less patient
+If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
 
@@
 Further, Reyna and Brainerd argued that people overweight numerators compared to denominators.
+The biases used when calculating ratios are relevant to intertemporal choices.
 
-Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
+Recent models of intertemporal choice have posited that people use ratios to decide between a smaller sooner and a larger later option [ITCH, DRIFT]. 
 An example intertemporal choice is choosing between \$5 in 1 day and \$6 in 5 days. 
@@
 
+We then apply our conceptual model to situations where temporal ratios are explicit.
+
 \section{Discounting Models}
@@
 Alternative based theories -- such as exponential and hyperbolic discounting -- posit that you calculate a discounted utility for each option.
-Whereas other models posit that people make attribute based comparisons meaning that they compare the money and time amounts to one another.
+Whereas other, attribute based models, posit that people make comparisons attribute based comparisons -- they compare the money and time amounts to one another.
 
@@
 
+The normative model of intertemporal choice is exponential discounting.
 The exponential discounting function is
 
-\begin{equation}\ref{eq:exp}
+\begin{equation}\label{eq:exp}
 	\delta_{exp} = \exp^{-rt}
@@
 
-Where $\delta$ is the discount fraction which each monetary amount is multiplied by and $r$ is the exponential discount rate. And the utlility of an option is:
+Where $\delta$ is the discount fraction which each monetary amount is multiplied by and $r$ is the exponential discount rate. And the utility of an option is:
 
-\begin{equation}\ref{eq:utilalt}
+\begin{equation}\label{eq:utilalt}
 	U(x, t) = x \delta
@@
 \subsection{ITCH}
+\label{ITCH}
 
@@
 
-\subsection{An example of TRM}
+\subsection{An example of TRaM}
 
@@
 A TRM decision maker would look at the two delays and calcluate their ratio. 
-Then they will use a rule of thumb to say how much relative compensation they need to account for that ratio. Below is an example of a TRM decision maker
-The time is 5 times as long therefore I will need 3 times as much money.
+Then they will use a rule of thumb to say how much relative compensation they need to account for that ratio.
+Below is an example of a TRM decision maker would make an intertemporal choice:
+
+The time is 5 times as long therefore I will need about 3 times as much money.
 
@@
 If we move both options up by five days then the ratio between times becomes $10/6$ = 1.66. A participant using the TRM heuristic would say that 
-The larger later time is about 1.5 times as long as the smaller sooner time therefore I need 1.25 times as much money.
+The larger later time is about 1.5 times as long as the smaller sooner time therefore I need about 1.25 times as much money.
+
+Further TRaM predicts that the salience of the temporal ratio means that it is more likely to be used. 
+For instance if the temporal ratio explicit, participants reliance on it would increase.
+
+\section{Combining ITCH, DRIFT, and TRaM }
+
+Read, Frederick, and Scholten (2013) show that when making intertemporal choices how the monetary amounts are framed alters patience. 
+They propose a Difference, Ratio, Interest, Finance, and Time (DRIFT) model of intertemporal choice. 
+For instance when an intertemporal choice is framed as \$100 now OR \$110 in 1 year people will rely on differences in monetary amounts.
+However if the choice is framed as \$100 now or an extra 10 \% after 1 year, people will rely more on ratios. 
+The DRIFT model of intertemporal choice shows that how monetary amounts are framed can alter levels of patience.
+
+Relatedly Read, Frederick, Orsel, and Rahman showed that how the times intertemporal choice are displayed can affect patience.
+Specifically Read and colleagues found that when times are described as dates (e.g.
+December 30th) people are more patient than when described as delays (e.g.
+in 1 month).
+Temporal framing can alter patience.
+
+Given that temporal framing can alter patience and that people rely on ratiosin intertemporal choice, we propose that making ratios between times explicit can alter patience. 
+For instance, an intertemporal choice between receiving \$10 in 2 days and \$11 in 12 days could be re-framed as \$10 in 2 days and \$11 in 6 times as long. 
+We proposed that this temporal framing leads to a greater weight being placed upon the time ratio term $\beta_{tR}$ and a lower weight on the time difference term $\beta_{tA}$ in equation \ref{eq:itch} .
+Taking the example and using the parameters in section \ref{ITCH} the probability of choosing larger later is $.26$.
+However if the ratio is weighted more heavily -- the absolute value of the beta coefficient goes up e.g.
+$\beta_{tR} = -.5 $ -- and the difference is weighted the same, eg.
+$\beta_{tA} = -.1$, the probability of choosing larger later is now $.0065$.
+
+Holding everything else constant by making the temporal ratio more salient, e.g.
+increasing the absolute value of $\beta_{tR}$,  monotonically decreases the probability of choosing larger later.
+We propose that making temporal ratios more salient increases their decision weight, leading to more impatient choices
+
+{\large [This CANNOT account for the change in the decision weight on temporal differences by making temporal ratios salient, which I am unsure of how to work out]. 
+}
+
+The TRaM conceptual model of intertemporal choice suggests that by making temporal ratios explicit that people's reliance on them will be higher.
+When people's reliance on temporal ratios is higher, the weight in which temporal ratios receive will be higher. 
+This weight has predictable consequences on how impatient participants are, which suggests ways in which firms can alter patience based on how times are presented.
 
@@
 In study 1 we will test this prediction of the degree of decreasing impatience depending on the change in ratio.
+We predict that when ratio changes are large, participants will show more diminishing impatience than when ratio changes are small.
 
 \subsection{Methods}
-Participants were either given 8 choices with either a small or large time ratio which was moved into the future by 5 days.
+Participants were in one of two conditions. 
+Six intertemporal choices were consistent across conditions and were used to calculate the participants discount fraction.
+The other two intertemporal choices varied across conditions, these choices varied in the extent to which the temporal ratio changed.
+We varied the ratio with either a small or large time ratio which was moved into the future by 5 days.
 
-In condition 1 participants were given 8 intertemporal choices however the focal questions were 
+In condition 1 participants the two focal questions were:  
 \$10 in 5 days or \$11 in 6 days
@@
 
-In condition 1 participants were given 8 intertemporal choices however the focal questions were 
-\$10 in 5 days or \$11 in 6 days
-and 
-\$10 in 10 days or \$11 in 11 days
-
-In condition 2 participants were given 8 intertemporal choices and the focal questions were 
+In condition 2 participants the two focal questions were: 
 \$10 in 5 days or \$11 in 15 days
@@
 
+In condition 1 the change in temporal ratio is small (1.2 to 1.1), whereas in condition 2 the change in temporal ratio is large (3 to 2).
+
 \subsubsection{Participants}
 
-\subsubsection{Results}
+We recruited [] Rutgers university undergraduates to complete the study for course credit.
+
+\subsection{Results}
 
@@
 
+\subsection{Discussion}
+
+Study 1 shows that the use of temporal ratios is more important than other cognitive models predicted. 
+If temporal ratios are important, by explicitly manipulating them, we should be able to alter levels of patience
+
+\section{Study 2}
+
+In study 2 we manipulate how easy it is to compare temporal ratios. 
+From ITCH and TRaM we hypothesize that participants who are shown explicit temporal ratios will make more impatient decisions
+
+\subsection{Methods}
+
+Each respondent answered 27 questions framed as intertemporal choices.
+We varied the time to the smaller sooner option ($SS_{time} =$ 5, 10, or 50 days), 
+the ratio between the smaller sooner and larger later times ($SS_{time} / LL_{time}  = $ 1.2, 2.2, or 3.2 times as long), 
+ the difference between the smaller sooner and larger later amount ($LL_{amnt} - SS_{amnt} =$ \$5, \$10, \$30). All smaller sooner amounts were \$47 [Or we could pull from a uniform distribution?].
+
+Each respondent was assigned to one of two frames: implicit temporal ratios and explicit temporal ratios. An example of both is given below
+
+\textbf{Implicit Temporal Ratio: } You can receive \$47 in 5 days or \$54 in 6 days
+
+\textbf{Explicit Temporal Ratio: } You can receive \$47 in 5 days or \$54 in 1.2 times as long
+
+\subsection{Results}
+
+We perform a multilevel logistic regression predicting choice with varying intercepts for each participant, the three within subject factors, and the between subject factor of temporal ratio frame.
+
+The results of the logistic regression are shown in table [].
+We see that participants in the explicit temporal ratio are more impatient frame than those in the implicit temporal ratio frame.
+[Is there an interesting interaction here e.g.
+we see this effect only when the ratio is large but not when it is small?]
+
+\section{Study 3}
+
+While study 2 showed that when time ratios are explicit, people are more impatient, it was framed as gains not losses.
+Most shipping situations involve a cost to the consumer. 
+In Study 3 we replicate study 2 but with losses.
+
+\subsection{Methods}
+
+The methods for study 3 are identical to study 2 except that all monetary amounts are framed as losses and not gains.
+
+\subsection{Results}
+
+We replicate the findings of study 2 when framed as losses.
+
+\section{Study 4}
+
+Studies 1 and 2 showed that explicit time ratios make people more impatient for both gains and losses, but do not specifically deal with a shipping scenario.
+Study 4 uses a hypothetical shipping scenario.
+
+\subsection{Methods}
+
+Participants were told to imagine that they had just decided to buy \$75 worth of merchandise on a popular on-line vendor. 
+[Should this strictly be between subjects?]
+
+They then were presented with two shipping options in one of the two frames.
+
+\textbf{Implicit Temporal Ratio}
+
+Option 1: Receive your order in two days for \$7
+
+Option 2: Receive your order in 10 days for \$3
+
+\textbf{Explicit Temporal Ratio}
+
+Option 1: Receive your order in two days for \$7
+
+Option 2: Receive your order in 5 times longer than option 1 for \$3
+
+\subsection{Results}
+
+We perform a logistic regression with shipping choice as the dependent measure and explicit vs implicit temporal ratio as the independent variable. 
+We find that participants are more likely to delay their shipping in the implicit temporal ratio condition.
+
+\subsection{Discussion}
+
+Study 4 shows that temporal ratios are an important driver of real world shipping choices. 
+By making temporal ratios explicit, people are more likely to be patient.
+
+\section{General Discussion}
+
+Across four studies we show temporal ratios are an important driver of temporal choice. 
+This importance has been understated by previous models of intertemporal choice, and by explicitly expressing temporal ratios people are more impatient.
+
 \end{document}

### 2015-11-30 shinyapps/itch_beta_tr.Rmd
updated version of paper with 4 studies

---
title: "ITCH_shiny"
author: "Daniel Wall"
date: "November 30, 2015"
output: html_document
runtime: shiny
---

```{r,echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
```

From @MarzilliEricson2015

According to ITCH the probability of choosing larger later is

$$P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)$$

Where $x^* = (x_1 + x_2)/2$

The function below calculates the probability that a participant will choose larger later.

```{r}
p_ll <- function(beta_intercept, beta_xa, beta_xr, beta_ta, beta_tr, x1, x2, t1, t2) {
  x_rat <- (x2 - x1)/ ((x2 + x1)/2)
  t_rat <- (t2 - t1)/ ((t2 + t1)/2)
  reg <- beta_intercept + beta_xa*(x2 - x1) + beta_xr*(x_rat) + 
    beta_ta*(t2 - t1) + beta_tr*(t_rat)
  plogis(reg)
} 
```

```{r, echo=FALSE}
inputPanel(
  sliderInput("beta_xa", "beta_xa",  min=.01, max = .99, step=.001, value= .1),
  sliderInput("beta_xr", "beta_xr",  min=.01, max = .99, step=.001, value= .1),
  sliderInput("beta_ta", "beta_ta",  min= -1, max = 0, step=.001, value= -.1),
  sliderInput("ssamnt", "ssamnt", value = 1,  min=0, max = 100, step = 1),
  sliderInput("sstime", "sstime",  value = 1,  min=0, max = 100, step = 1),
  sliderInput("llamnt", "llamnt", value = 2,  min=0, max = 100, step = 1),
  sliderInput("lltime", "lltime",  value = 10,  min=0, max = 100, step = 1)
)

renderPlot({
  itch_val_df <- data.frame(beta_tr = seq(-1, 0, by = .001))
  itch_val_df$beta_xa <- input$beta_xa
  itch_val_df$beta_xr <- input$beta_xr
  itch_val_df$beta_ta <- input$beta_ta
  itch_val_df$ssamnt <- input$ssamnt  
  itch_val_df$sstime <- input$sstime
  itch_val_df$llamnt <- input$llamnt
  itch_val_df$lltime <- input$lltime

  itch_val_df <- mutate(itch_val_df, p_ll = p_ll(0.0, beta_xa, beta_xr, beta_ta, beta_tr, ssamnt, llamnt, sstime, lltime))

  ggplot(itch_val_df, aes(x = beta_tr, y = p_ll, group = 1)) +
    geom_line() +
    xlab("beta_tr") +
    ylab("probability of choosing larger later")
})

```

wireframe

```{r}
inputPanel(
  sliderInput("beta_xa2", "beta_xa2",  min=.01, max = .99, step=.001, value= .1),
  sliderInput("beta_xr2", "beta_xr2",  min=.01, max = .99, step=.001, value= .1),
  sliderInput("ssamnt2", "ssamnt2", value = 1,  min=0, max = 100, step = 1),
  sliderInput("sstime2", "sstime2",  value = 1,  min=0, max = 100, step = 1),
  sliderInput("llamnt2", "llamnt2", value = 2,  min=0, max = 100, step = 1),
  sliderInput("lltime2", "lltime2",  value = 10,  min=0, max = 100, step = 1)
)

renderPlot({
  itch_val_df2 <- data.frame(beta_tr = seq(-1, 0, by = .001),
                            beta_ta = seq(-1, 0, by = .001))

  itch_val_df2$beta_xa <- input$beta_xa2
  itch_val_df2$beta_xr <- input$beta_xr2
  itch_val_df2$ssamnt <- input$ssamnt2  
  itch_val_df2$sstime <- input$sstime2
  itch_val_df2$llamnt <- input$llamnt2
  itch_val_df2$lltime <- input$lltime2

  itch_val_df2 <- mutate(itch_val_df2, p_ll = p_ll(0.0, beta_xa, beta_xr, beta_ta, beta_tr, ssamnt, llamnt, sstime, lltime))

  ggplot(itch_val_df, aes(x = beta_tr, y = p_ll, group = 1)) +
    geom_line() +
    xlab("beta_tr") +
    ylab("probability of choosing larger later")
})
```

### 2015-12-14 wave_10dec/0_gather_clean_data.R
Analyze data for ashwani tomorrow.

@@ -1,0 +1,106 @@
+#--------------
+# Gather and clean data for the titrated values for ITC uncertainty
+# Daniel Wall
+# 3 November 2015
+#--------------
+
+setwd("/git_repositories/timeratios/wave_10dec/")
+source("load_packages.R")
+rm(list = ls())
+
+
+#----read in the data----
+
+timerat <- read_csv("raw_data/Time_Ratios_6_December.csv")
+
+
+# change column names
+names(timerat) <- tolower(names(timerat))
+
+# remove test runs
+timerat <- timerat %>%
+  filter(finished == 1)
+
+
+
+
+# get data so I can work on it for when people are switching
+# dissimilarity ids
+diss_id <- read.csv("raw_data/diss_ids.csv") 
+
+# filter data 
+diss <- timerat[, str_detect(names(timerat), "^(responseid|ul)")] %>%
+  melt(id = "responseid") %>%
+  filter(!(is.na(value))) %>%
+  group_by(responseid)
+
+# get which condition the participant was in
+diss_m <- diss %>%
+  group_by(responseid) %>%
+  summarise(ratcng = ifelse("ul5_6_1" %in% variable, "small", "large"))
+
+diss <- merge(diss, diss_m, by = "responseid", all.x=TRUE) %>%
+  mutate(ss_time = str_extract(variable, "\\d+"),
+         ll_time = str_extract(variable, "\\d+_1$") %>% str_extract("\\d+"))
+
+
+# now merge with dissimilarity ids
+diss <- merge(diss, diss_id, by = c("ss_time", "ll_time"))
+
+dissc <- diss %>%
+  dcast(responseid + id + ratcng ~ delayed, value.var = "value") %>%
+  mutate(diss_diff = delayed - not_delayed,
+         ratcng = factor(ratcng))
+
+disssum <- dissc %>%
+  group_by(ratcng) %>%
+  summarise(mean_diffcng = mean(diss_diff, na.rm = TRUE),
+            se_diffcng = std.error(diss_diff, na.rm = TRUE)) %>%
+  mutate(lower = mean_diffcng - 1.96*se_diffcng,
+         upper = mean_diffcng + 1.96*se_diffcng)
+
+ggplot(disssum, aes(x = ratcng, y = mean_diffcng)) +
+  geom_point() +
+  geom_errorbar(aes(ymin = lower, ymax = upper))
+  
[…]

### 2015-12-14 wave_10dec/load_packages.R

@@ -1,0 +1,17 @@
+#--------------
+# Load packages
+# Daniel Wall
+# 3 November 2015
+#--------------
+library(plyr)
+library(dplyr)
+library(stringr)
+library(readr)
+library(reshape2)
+library(broom)
+library(plotrix)
+library(ggplot2)
+library(lme4)
+library(effects)
+library(car)
+library(broom)

### 2015-12-14 paper/paperintro.tex
some new introduction stuffs

@@
 
+%\usepackage{cogsci}
+\usepackage{pslatex}
+\usepackage{apacite}
+\usepackage{graphicx}
+\usepackage{booktabs,caption,fixltx2e}
 %opening
@@
 \begin{abstract}
-	Choices involving time, intertemporal choices, are ubiquitous -- for instance choosing a shipping speed for your latest Amazon order. 
-	Further, people use ratios in everyday life -- intertemporal choices are no different, 
+	When selecting our shipping speed on Amazon we are faced with a choice between getting our package sooner but paying more money, or getting our package later and paying less money. 
+	This phenomenon -- known as intertemporal choice is the subject of this paper.
+%	Choices involving time, intertemporal choices, are ubiquitous -- for instance choosing a shipping speed for your latest Amazon order. 
+	People use ratios in everyday life -- intertemporal choices are no different, 
 	recent cognitive models of intertemporal choice show that people calculate ratios of both money and time. 
@@
 
-Many of our daily choices involve making decisions involving time in some respect, often taking the form of a the tradeoff between receiving something of smaller value now and receiving something of larger value in the future.
-For instance when a consumer selects a shipping speed when shopping online they are making an intertemporal choice.
-These decisions require consumers to, perhaps implicitly, calculate how much waiting for the slower shipping speed is worth.
+All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible. 
+Sometimes that itch may lead to us choosing to expedite our shipping at considerable cost. 
+There are subtle ways in which the presentation of time can alter patience. 
+For instance people are more impatient when delays are calendar dates as opposed to delays into the future \cite{Read2005}. 
+Marketers have used temporal frames in shipping scenarios.
+For example as opposed to stating that you will receive your package in 2 weeks, Amazon sometimes states that you receive your package on January 23rd. 
+As per \citeA{Read2005} this frame will increase patience.
 
-Firms could have differing preferences on whether a consumer is impatient or patient when choosing a shipping speed. 
+%Many of our daily choices involve making decisions involving time in some respect, often taking the form of a the tradeoff between receiving something of smaller value now and receiving something of larger value in the future.
+%For instance when a consumer selects a shipping speed when shopping online they are making an intertemporal choice.
+%These decisions require consumers to, perhaps implicitly, calculate how much waiting for the slower shipping speed is worth.
+
+Firms could, however, prefer that a consumer is impatient when choosing a shipping speed. 
 If a firm's inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster.
 It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
-This understanding how to optimally display information leads to novel predictions for how to display shipping speeds.
+This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds.
 
@@
 
+\section{Study}
+
+\subsection{Methods}
+
+\subsubsection{Participants}
+
+Ninety three Rutgers students took part in this study as a part of their course requirements.
+
+\subsubsection{Design}
+
+Participants were randomly assigned to one of two conditions -- small ratios between times and large ratios between times.  
+As seen in Table~\ref{tab:stimuli}, 24 unique intertemporal choices were created. 
+The first column shows the first two times, the third column shows the delay added to both of the first two columns which yeild columns 4 and 5. 
+The $SS_{amnt}$ and $LL_{amnt}$ columns were chosen using intuition to generate half of the participants choosing the smaller sooner option and half choosing the larger later option. 
+The last column indicates the condition which participants saw, e.g. small ratio versus large ratio change
+
+\begin{table}[!htbp] \centering 
+	\caption{Stimuli Used for Study \_ } \label{tab:stimuli}
+\begin{tabular}{ | l | l | l | l | l | l | l | l | }
+	\hline
+	$SS_{time}$ & $LL_{time}$ & delay & $SS_{time}^{delay}$ & $LL_{time}^{delay}$ & $SS_{amnt}$ & $LL_{amnt}$ & Ratio Change \\ \hline
+	5 & 6 & 5 & 10 & 11 & 105 & 106 & Small  \\ \hline
+	7 & 9 & 17 & 24 & 26 & 122 & 127 & Small  \\ \hline
+	10 & 14 & 14 & 24 & 28 & 144 & 149 & Small  \\ \hline
+	1 & 2 & 5 & 6 & 7 & 170 & 171 & Small  \\ \hline
+	20 & 25 & 22 & 42 & 47 & 108 & 114 & Small  \\ \hline
+	15 & 20 & 25 & 40 & 45 & 140 & 144 & Small  \\ \hline
+	5 & 15 & 5 & 10 & 20 & 105 & 106 & Large \\ \hline
+	7 & 21 & 17 & 24 & 38 & 122 & 127 & Large \\ \hline
+	10 & 22 & 14 & 24 & 36 & 144 & 149 & Large \\ \hline
+	1 & 10 & 5 & 6 & 15 & 170 & 171 & Large \\ \hline
+	20 & 40 & 22 & 42 & 62 & 108 & 114 & Large \\ \hline
+	15 & 75 & 25 & 40 & 100 & 140 & 144 & Large \\ \hline
+\end{tabular}
+\end{table}
+
+\subsubsection{Procedure}
+
+Participants initially granted informed consent. 
+Then were randomly selected to be in one of the two conditions Small or Large ratio change. 
+Participants then made ``unlikeness'' judgments between each set of times in their condition. 
+Furhter, non-delayed and delayed times were blocked together the order in which the questions were asked and the order in which the blocks were presented was randomized. 
+The blocking of items was to minimize the likelihood that participants saw items in the same row of Table~\ref{tab:stimuli} consecutively. 
+For instance a participant in the small ratio change condition answered how unlike each  the 6 non-delayed times were then answered the how unlike eachother each of the 6 delayed times were.
+
+Participants then completed an unrelated task about buying cars, which will not be reported further.
+
+Next particpants answered each of the intertemporal choices in theri condition. 
+The intertemporal choices were blocked similarly to the unlikeness judgments. 
+Next participants were asked the extent to which they used ratios and the extent to which they used differences when making intertemporal choices. 
+Finally participants answered demographics questions.
+
+\subsection{Results}
+
 \section{General Discussion}
@@
 
+\section{Limitations and Future Directions}
+
+\bibliographystyle{apacite}
+
+\setlength{\bibleftmargin}{.125in}
+\setlength{\bibindent}{-\bibleftmargin}
+
+\bibliography{C://git_repositories/intertemporal_uncertainty/paper/library}
+
 \end{document}

### 2015-12-14 wave_10dec/presentation/methods_results_14dec.Rmd
Present data for 10 December Wave

---
title: "Methods and Results 14 December 2015"
author: "Daniel Wall"
date: "December 14, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}

setwd("/git_repositories/timeratios/wave_10dec/")
source("load_packages.R")
rm(list = ls())

load("analyzed_data/analyzed_data.Rdata")
# opts_chunk$set(echo=FALSE) 
```

# Summary

We are trying to see if diminishing impatience increases to a greater exetent when ratios change a lot.

# Unlikeness scores

We want to see a bigger change in perceived differences between items when the ratio change is large as compared to small.

I subtracted the not delayed onset from the delayed onset.

## Mean and CI of differences

```{r, echo=FALSE}
ggplot(disssum, aes(x = ratcng, y = mean_diffcng)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  ylab("Difference between not delayed and delayed intervals")

```

## Model predicting change in unlikeness

```{r, echo=FALSE}
Anova(diss_lmer, type = 3)
summary(diss_lmer)
```

### Plot of effects

```{r, echo=FALSE}
ggplot(eff_df, aes(x = ratcng, y = fit)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper))
```

# Intertemporal choices

We want there to be a large ratio change in diminishing impatience for the large ratio change as compared to the small ratio change.

## Graph of mean percent larger later

```{r, echo=FALSE}

ggplot(itc_sum,
       aes(x = diss_set, y = estimate, group = delayed, fill = delayed, 
           ymax = conf.high,  ymin = conf.low)) +
  geom_bar(stat="identity", position = "dodge") +
  ylab("Percent Larger Later") +
  geom_errorbar(position = "dodge")

```

## Logistic regression predicting choices

```{r, echo=FALSE}
Anova(itc_lmer, type=3)
summary(itc_lmer)
```

### Effects plot from logistic regression

```{r, echo=FALSE}
ggplot(itc_eff, aes(x = diss_set, y = fit, group = delayed,
                    fill = delayed,
                    ymin = lower, ymax = upper)) +
  geom_bar(stat="identity", position = "dodge") +
  geom_errorbar(position = "dodge") +
  ylab("Logistic Regression Coefficient")

```

### 2015-12-14 wave_10dec/0_gather_clean_data.R

@@ -61,7 +61,8 @@ disssum <- dissc %>%
 
 ggplot(disssum, aes(x = ratcng, y = mean_diffcng)) +
   geom_point() +
-  geom_errorbar(aes(ymin = lower, ymax = upper))
+  geom_errorbar(aes(ymin = lower, ymax = upper)) +
+  ylab("Difference between not delayed and delayed intervals")
   
 # mixed effects model
 diss_lmer <- lmer(diss_diff ~ ratcng + (1|responseid),
@@ -87,20 +88,48 @@ itc <- timerat[, str_detect(names(timerat), "^(responseid|itc\\d+_\\d+)")] %>%
   group_by(responseid) %>%
     merge(diss_id, by = c("ss_time", "ll_time"))
   
-  
+#summarise and plot the data
 itc_sum <- itc %>%
   group_by(diss_set, delayed) %>%
   summarise(n_ll = sum(value == "2"),
             n_tot = n()) %>%
   mutate(per_ll = n_ll/n_tot) %>%
   group_by(diss_set, delayed) %>%
-  do( tidy(prop.test(.$n_ll, .$n_tot)))
+  do( tidy(prop.test(.$n_ll, .$n_tot))) %>%
+  ungroup %>%
+  mutate(diss_set = factor(diss_set),
+         delayed = factor(delayed))
 
 
 ggplot(itc_sum,
-       aes(x = diss_set, y = estimate, group = delayed, ymax = conf.high,  ymin = conf.low)) +
+       aes(x = diss_set, y = estimate, group = delayed, fill = delayed, 
+           ymax = conf.high,  ymin = conf.low)) +
   geom_bar(stat="identity", position = "dodge") +
+  ylab("Percent Larger Later") +
   geom_errorbar(position = "dodge")
   
+
+# now lets run a model on this
+itc <- itc %>%
+  mutate(value = factor(value),
+         diss_set = factor(diss_set),
+         delayed = factor(delayed))
+
+itc_lmer <- glmer(value ~ diss_set * delayed + (1|responseid),
+                  itc, family = "binomial")
+
+summary(itc_lmer)
+
+itc_eff <- Effect(c("diss_set",  "delayed"), itc_lmer) %>%
+  as.data.frame()
+
+ggplot(itc_eff, aes(x = diss_set, y = fit, group = delayed,
+                    fill = delayed,
+                    ymin = lower, ymax = upper)) +
[…]

### 2015-12-15 paper/paperintro.tex
Write up intro and beginning of study 1

@@
 %opening
-\title{An exploration of how ratios affect time discounting}
+\title{TRaM: Analyzing temporal framing in Intertemporal Choice}
 \author{Daniel Wall}
@@
 \begin{abstract}
-	When selecting our shipping speed on Amazon we are faced with a choice between getting our package sooner but paying more money, or getting our package later and paying less money. 
-	This phenomenon -- known as intertemporal choice is the subject of this paper.
-%	Choices involving time, intertemporal choices, are ubiquitous -- for instance choosing a shipping speed for your latest Amazon order. 
-	People use ratios in everyday life -- intertemporal choices are no different, 
-	recent cognitive models of intertemporal choice show that people calculate ratios of both money and time. 
-	These models account for diminishing impatience -- when items are moved into the future, people are more patient. 
-	Our goal is twofold: 1) to show that previous models of intertemporal choice do not properly account for temporal ratios and present a conceptual model of time discounting, and  2) by explicitly manipulating how people use ratios of time, we can alter how people make intertemporal choices.
-	This leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
-	Further we show that in a hypothetical shipping scenario that when temporal ratios are explicit, people are more impatient.
+	When selecting a shipping speed on Amazon we are faced with a choice between getting our package sooner but paying more money, or getting our package later and paying less money. 
+	The general phenomenon underlying our expedited shipping is impatience, we want good things immediately and we must be compensated for waiting. 
+	Previous research on intertemporal framing has shown that displaying times as dates -- January 23rd -- as opposed to delays -- 6 weeks -- increases patience. 
+	Other research has shown that people use both differences and ratios of time in intertemporal choices. 
+%	These two results suggest that framing times as ratios alters patience. 
+%	In three experiments I manipulate the temporal presentation from the standard -- 1 day versus 5 days --  to making the ratio explicit -- 1 day or 5 times as long.  
+	Across three experiments I show that making temporal ratios explict leads to differing levels to increase impatience in standard  intertemporal choices framed as gains, intertemporal choices framed as losses, and for an explicit shipping scenario.
+	Taken together these results suggest that framing delays as ratios leads to greater impatience.
+
+%
+%	People use ratios in everyday life -- intertemporal choices are no different, 
+%	recent cognitive models of intertemporal choice show that people calculate ratios of both money and time. 
+%	These models account for diminishing impatience -- when items are moved into the future, people are more patient. 
+%	Our goal is twofold: 1) to show that previous models of intertemporal choice do not properly account for temporal ratios and present a conceptual model of time discounting, and  2) by explicitly manipulating how people use ratios of time, we can alter how people make intertemporal choices.
+%	This leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
+%	Further we show that in a hypothetical shipping scenario that when temporal ratios are explicit, people are more impatient.
 
@@
 
-All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible. 
-Sometimes that itch may lead to us choosing to expedite our shipping at considerable cost. 
-There are subtle ways in which the presentation of time can alter patience. 
-For instance people are more impatient when delays are calendar dates as opposed to delays into the future \cite{Read2005}. 
-Marketers have used temporal frames in shipping scenarios.
-For example as opposed to stating that you will receive your package in 2 weeks, Amazon sometimes states that you receive your package on January 23rd. 
-As per \citeA{Read2005} this frame will increase patience.
+All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible; sometimes leading us choosing expedited shipping, even at considerable cost. 
+Generally on Amazon we see shipping times as the number of days until you'll receive your package, e.g.  Receive your package in 2 days. 
+However, an identical choice to the one above would be Recieve your package on December 16th. 
+\citeA{Read2005} showed that people are more impatient when times are framed as days as compared framed as to dates. 
+Accordingly when shopping online delays are occasionally framed as dates.
+
+However a firm may want consumers to be impatient and choose the expedited shipping option. 
+If a firm's inventory is high they may want to nudge consumers to ship their items faster. 
+It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
+This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds.
+
+How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
+Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
+%For instance \citeA{Read2013} found that by making ratios in the 
+However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
+Across 3 studies we show that by making ratios explicit people are more impatient in intertemporal choices. 
+%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
+This result holds for intertemporal choices framed as gains, losses, and in an explict shipping scenario. 
+%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
+
+%
+%There are subtle ways in which the presentation of time can alter patience. 
+%For instance people are more impatient when delays are calendar dates as opposed to delays into the future \cite{Read2005}. 
+%Marketers have used temporal frames in shipping scenarios.
+%For example as opposed to stating that you will receive your package in 2 weeks, Amazon sometimes states that you receive your package on January 23rd. 
+%As per \citeA{Read2005} this frame will increase patience.
 
@@
 
-Firms could, however, prefer that a consumer is impatient when choosing a shipping speed. 
-If a firm's inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster.
-It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
-This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds.
+%Firms could, however, prefer that a consumer is impatient when choosing a shipping speed. 
+%If a firm's inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster.
+%It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
+%This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds. 
+%
+%Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
+%Across [] studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
+%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios. 
+%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
 
-Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
-Across [] studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
-Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios. 
-If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
+\section{Intertemporal Choices}
 
-\section{Ratios and Intertemporal Choices}
+Intertemporal choices in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}.
+
+There are multiple categories of research on these intertemporal choices (see \citeA{Read2013} for an overview) however the current paper belongs to research which investigates how impatience is altered by option frames.
+An example of option framing in intertemporal choice is the hidden zero effect -- when the implicit zeros in intertemporal choices are made explicit people are more patient \cite{Magen2008}. 
+For instance when a choice between ``receiving \$100 now'' and ``\$110 in 6 months'' is reframed to be  ``receiving \$100 now \textit{and nothing in six months}'' and ``\$110 in 6 months \textit{and nothing now}'' people are more patient.
+
+Additionally \citeA{Read2013} show that making the ratio between amounts salient affects intertemporal choices. 
+Specifically when the magnitude of amounts is small, reframing the amounts as interest rates leads to increased patience.
+Taking the above example, when reframed as an interest rate, e.g.
+a choice between  ``receive \$100 now'' and ``receive 10\% more in six months'' people are more patient.
+\citeA{Read2013} posit that intrest framed amounts lead to an increased reliance on ratios.
+
+Other work by \citeA{Read2005} shows that how a times are framed in intertemporal choices can alter patience. 
+Specifically they find that by reframing the delay above to be a date induces greater patience. 
+Taking the above example, people are more patient when reframed it as a choice between  ``receive \$100 now'' and ``receive \$110 on June 15''.
+This demonstrates that changing how times are framed can affect patience.
+
+In other work on intertemporal choices, \citeA{MarzilliEricson2015} show that in  standard framed intertemporal choices, people rely on both differences and ratios between both monetary amounts and times.
+ Their model -- Intertemporal Choice Heuristic (ITCH) -- does better at predicting out of sample intertemporal choices than classic models of intertemporal choices such as hyperbolic or exponential disounting.
+As seen in Equation~\ref{eq:itch}  the functional form of the itch model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting.
+
+[What are they?]
+
+ \begin{equation}\label{eq:itch}
+ P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
+ \end{equation}
+
+This paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience.
+Specifically, I focus on how framing times as ratios can decrease patience.
+When delays are sufficiently small, and temporal ratios are made explicit, our attention will be focused on how much \textit{longer} we will wait. 
+This focus on how much longer we will wait will lead to decreased patience when times are framed as ratios.
+
+I operationalize this prediction across three studies.
+Study 1 investigates how making temporal ratios explicit in intertemporal choices leads to greater impatience.
+
+Study 2 extends the results of Study 1 to losses. 
+Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
+For instance deciding to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
+Intertemporal choices framed as losses take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
+Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
+Study 2 investigates how framing times as ratios leads to decreased patience in losses.
+
+Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
+Shipping scenarios are displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
+Although choosing a shipping speed is similar to the intertemporal choices there is a subtle, yet possibly significant difference between the two.
+Shipping decisions are paying more now to receive an item sooner whereas the intertemporal choices shown in study 2 are paying something smaller now to avoid paying something larger later.
+
+\subsection{Hypothesis}
+
+\subsubsection{Hypothesis 1}
+According to the ITCH model people rely on ratios and differences in temporal choices. 
+A prediction of the model is that the higher the weight given to the temporal ratio the more impatient participants will be.
+
+\textit{Hypothesis 1} In gain intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.
+
+\subsubsection{Hypothesis 2}
+Similar to hypothesis 1 when framed as losses making temporal ratios explicit people will be more impatient.
+
+\textit{Hypothesis 2} In loss intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.
+
+\subsubsection{Hypothesis 3}
+Similar to hypothesis 2 when framed as a shipping scenario making temporal ratios explicit people will be more impatient.
+
+\textit{Hypothesis 3} In shipping scenarios, when times are expressed as ratios, people will be more impatient than when expressed as delays.
+
+\section{Study 1}
+In study 1 participants chose between smaller sooner and larger later options. 
+These options were displayed in one of two frames: implicit temporal ratios and explicit temporal ratios. 
+Study 1 tests Hypothesis 1 above
+
+\subsection{Method}
+Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
+Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
+Furhter, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study.
+
+Participants completed informed consent, then were given a test to ensure they understood the paradigm.
+They then were assigned to one of the two conditions: implicit ratio or explicit ratio. 
+They then were shown the intertemporal choices in Table []. 
+Each intertemporal choice was shown in the following format.
+
+Insert a qualtrics picture here.
+
+Choices were displayed in a random order to each participant.
+
+After completeing their choices participants completed demographic information.
+
+\subsection{Results}
+
+The dependent variable in this experiment was the number of times a participant chose the larger later -- e.g. more patient -- reward.
+
+each participant chose the larger-later reward ?LL?.As Table 2 shows, there was a strong date/delay effect: Twice as many people chose LL in the Date condition than in the Week or Month conditions.
+Differences across conditions were tested formally with Kruskal- Wallis tests.
+An overall test confirmed a signifi- cant difference between the conditions (?2?2?
+= 21?1,
 
@@
 
-\begin{equation}\label{eq:itch}
-	P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
-\end{equation}
+%\begin{equation}\label{eq:itch}
+%	P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
+%\end{equation}
 

### 2015-12-15 paper/paperintro.tex
Start making intro and literature review better

@@
 %opening
-\title{TRaM: Analyzing temporal framing in Intertemporal Choice}
+\title{To Ship or Not to Ship: Explict Temporal Ratios Lead to Impatience}
 \author{Daniel Wall}
@@
 All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible; sometimes leading us choosing expedited shipping, even at considerable cost. 
-Generally on Amazon we see shipping times as the number of days until you'll receive your package, e.g.  Receive your package in 2 days. 
-However, an identical choice to the one above would be Recieve your package on December 16th. 
-\citeA{Read2005} showed that people are more impatient when times are framed as days as compared framed as to dates. 
-Accordingly when shopping online delays are occasionally framed as dates.
+Generally on Amazon we see shipping times as the number of days until you will receive your package, e.g.  ``Receive your package in 2 days''. 
+The shipping time could also be described as ``Receive your package on December 17th''. 
+This reframing of delays -- 2 days -- into dates -- December 17th -- has been shown to increase patience \cite{Read2005}.
+Occasionally Amazon frames shipping times as dates, possibly to induce customers to be more patient and receive their items later.
 
@@
 However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
-Across 3 studies we show that by making ratios explicit people are more impatient in intertemporal choices. 
+Across 3 studies I show that by making ratios explicit people are more impatient in intertemporal choices. 
 %Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
-This result holds for intertemporal choices framed as gains, losses, and in an explict shipping scenario. 
+This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
 %If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
@@
 
-\section{Intertemporal Choices}
+\section{Literature Review}
 
-Intertemporal choices in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}.
+Intertemporal preferences measured in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}. 
+There are four general categories of research on  intertemporal choices (see \citeA{Read2013} for an overview).
+The first focuses on how magnitude of both time and money effects patience \cite{Scholten2006, Thaler1981}
+ the current paper belongs to research investigating how patience is altered by option frames. 
+ The second stream focuses on how choices relate to real world behavior.
+ Individual differences in imputed discount functions predict real world phenomenon such as credit card debt,  mortgage choice, and smoking \cite{MacKillop2011, Meier2009, Johnson2011}. 
+ The third stream focuses on how emotional states can alter intertmporal preferences, e.g. being seeing freshly baked bread \cite{Li2008}. 
+ The fourth stream investigates how the framing of options affects patience.
 
-There are multiple categories of research on these intertemporal choices (see \citeA{Read2013} for an overview) however the current paper belongs to research which investigates how impatience is altered by option frames.
 An example of option framing in intertemporal choice is the hidden zero effect -- when the implicit zeros in intertemporal choices are made explicit people are more patient \cite{Magen2008}. 
@@
 a choice between  ``receive \$100 now'' and ``receive 10\% more in six months'' people are more patient.
-\citeA{Read2013} posit that intrest framed amounts lead to an increased reliance on ratios.
+\citeA{Read2013} posit that interest framed amounts lead to an increased reliance on ratios of monetary amounts.
 
-Other work by \citeA{Read2005} shows that how a times are framed in intertemporal choices can alter patience. 
-Specifically they find that by reframing the delay above to be a date induces greater patience. 
-Taking the above example, people are more patient when reframed it as a choice between  ``receive \$100 now'' and ``receive \$110 on June 15''.
-This demonstrates that changing how times are framed can affect patience.
+Beyond amounts, times can also be reframed.
+The most famous example of temporal reframing is  Lincoln's Gettysburg Address which made 87 years into the ``Four score and seven years ago''.
+\citeA{Read2005} shows that how a times are framed in intertemporal choices can alter patience. 
+Taking the above example, people are more patient when reframed it as a choice between  ``receive \$100 now'' and ``receive \$110 on June 15'' -- this is known as the delay/date effect. 
+The delay/date effect demonstrates that perceived temporal distance is dependent on how delays are framed.
 
-In other work on intertemporal choices, \citeA{MarzilliEricson2015} show that in  standard framed intertemporal choices, people rely on both differences and ratios between both monetary amounts and times.
- Their model -- Intertemporal Choice Heuristic (ITCH) -- does better at predicting out of sample intertemporal choices than classic models of intertemporal choices such as hyperbolic or exponential disounting.
-As seen in Equation~\ref{eq:itch}  the functional form of the itch model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting.
+In other work on intertemporal choices, \citeA{MarzilliEricson2015} show that in intertemporal choices framed as ``receiving \$100 now'' and ``\$110 in 6 months'', people rely on both differences and ratios of the  attributes of intertemporal choices -- monetary amounts and times.
+Their model -- the Intertemporal Choice Heuristic (ITCH) -- does better at predicting intertemporal choices than classic models, such as hyperbolic or exponential discounting. 
+As seen in Equation~\ref{eq:itch}  the functional form of the ITCH model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting.
 
@@
 
-This paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience.
+This work aims to demonstrate that the fourth research stream -- that patience is affected by how options are framed -- can affect real world phenomenon.
+Specifically, this paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience.
 Specifically, I focus on how framing times as ratios can decrease patience.
-When delays are sufficiently small, and temporal ratios are made explicit, our attention will be focused on how much \textit{longer} we will wait. 
-This focus on how much longer we will wait will lead to decreased patience when times are framed as ratios.
-
-I operationalize this prediction across three studies.
-Study 1 investigates how making temporal ratios explicit in intertemporal choices leads to greater impatience.
-
-Study 2 extends the results of Study 1 to losses. 
-Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
-For instance deciding to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
-Intertemporal choices framed as losses take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
-Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
-Study 2 investigates how framing times as ratios leads to decreased patience in losses.
-
-Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
-Shipping scenarios are displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
-Although choosing a shipping speed is similar to the intertemporal choices there is a subtle, yet possibly significant difference between the two.
-Shipping decisions are paying more now to receive an item sooner whereas the intertemporal choices shown in study 2 are paying something smaller now to avoid paying something larger later.
+When delays are sufficiently small, and when temporal ratios are made explicit, our attention will be focused on the ratio of how much \textit{longer} we will wait. 
+For instance reframing a choice between ``receiving \$100  in 1 day'' and ``\$110 in 10 days'' to being a choice between ``receiving \$100  in 1 day'' and ``\$110 in 10 \textit{times as long}'', will focus participants on how much longer they will wait for \$110 as opposed to receiving the \$100 today.
+This increased focus on how long they will wait, leads to decreased patience when times are framed as ratios.
 
@@
 
+I test these hypotheses across three studies.
+Study 1 investigates Hypothesis 1 by making temporal ratios explicit in intertemporal.
+
+Study 2 extends the results of Study 1 to losses. 
+Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
+For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
+Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
+Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
+Study 2 investigates how framing times as ratios leads to decreased patience in losses.
+
+Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
+Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
+Although choosing a shipping speed is similar to the previous intertemporal choices there is a subtle, yet possibly significant difference between the two.
+Shipping decisions can be thought of as paying more now to receive an item sooner whereas the intertemporal choices shown in Study 2 can be thought of as paying something smaller now to avoid paying something larger later.
+
 \section{Study 1}
@@
 Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
-Furhter, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study.
+Further, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
+Additionally only participants who passed an attention check were included in the final analysis.
+These precautions were taken to ensure that participants completed
 
@@
 
-Insert a qualtrics picture here.
+Insert a Qualtrics picture here.
 
@@
 
-After completeing their choices participants completed demographic information.
+After completing their choices participants completed demographic information.
 

### 2015-12-15 paper/paperintro.tex
Now with properly formatted table

@@
 \usepackage{booktabs,caption,fixltx2e}
+\usepackage{rotating}
+\usepackage[letterpaper, margin=1in]{geometry}
+
+\usepackage{pdflscape}
+
 %opening
@@
 Specifically, this paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience.
+
 Specifically, I focus on how framing times as ratios can decrease patience.
@@
 
-\textit{Hypothesis 1} In gain intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.
+\textit{Hypothesis 1:} In gain intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.
 
@@
 
-\textit{Hypothesis 2} In loss intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.
+\textit{Hypothesis 2:} In loss intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.
 
@@
 
-\textit{Hypothesis 3} In shipping scenarios, when times are expressed as ratios, people will be more impatient than when expressed as delays.
+\textit{Hypothesis 3:} In shipping scenarios, when times are expressed as ratios, people will be more impatient than when expressed as delays.
 
 I test these hypotheses across three studies.
-Study 1 investigates Hypothesis 1 by making temporal ratios explicit in intertemporal.
+Study 1 tests Hypothesis 1 by making temporal ratios explicit in intertemporal choices.
 
@@
 These options were displayed in one of two frames: implicit temporal ratios and explicit temporal ratios. 
-Study 1 tests Hypothesis 1 above
+Study 1 tests Hypothesis 1: In gain intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.
 
@@
 Each intertemporal choice was shown in the following format.
+
+\begin{landscape}
+\begin{table}[!ht]
+	\caption{Amounts and times of Stimuli for Studies 1-3} \label{tab:stimuli}
+	\includegraphics[]{Stimuli_For_Study.pdf}
+\end{table}
+\end{landscape}
 

### 2015-12-15 paper/paperintro.tex
now with study 1 almost done

@@
 %opening
-\title{To Ship or Not to Ship: Explict Temporal Ratios Lead to Impatience}
+\title{Delay d'ETRE: Explict Temporal Ratios Lead to Impatience}
 \author{Daniel Wall}
@@
 However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
+Further although shipping scenarios are a form of intertemporal choice, they differ from how most studies operationalize intertemporal choices. 
 Across 3 studies I show that by making ratios explicit people are more impatient in intertemporal choices. 
@@
 This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
+As opposed to other studies which showed that reframed times as dates increases patience, framing times as ratios decreases patience.
+Firms wanting to decrease patience in their consumers may want to frame intertemporal choices as ratios as opposed to delays.
+
 %If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
@@
 Individual differences in imputed discount functions predict real world phenomenon such as credit card debt,  mortgage choice, and smoking \cite{MacKillop2011, Meier2009, Johnson2011}. 
- The third stream focuses on how emotional states can alter intertmporal preferences, e.g. being seeing freshly baked bread \cite{Li2008}. 
+The third stream focuses on how emotional states can alter intertmporal preferences, e.g. being shown freshly baked bread \cite{Li2008}. 
 The fourth stream investigates how the framing of options affects patience.
@@
 \subsubsection{Hypothesis 1}
-According to the ITCH model people rely on ratios and differences in temporal choices. 
-A prediction of the model is that the higher the weight given to the temporal ratio the more impatient participants will be.
+Given that people rely on differences and ratios between amounts and times in intertemporal choice, increasing the reliance on ratios for times with relatively small differences should lead to decreased patience.
 
@@
 \subsubsection{Hypothesis 3}
-Similar to hypothesis 2 when framed as a shipping scenario making temporal ratios explicit people will be more impatient.
+Related to hypothesis 2 when framed an intertemporal chioce is framed as a shipping scenario,  making temporal ratios explicit people will induce less patience.
 
@@
 
-I test these hypotheses across three studies.
-Study 1 tests Hypothesis 1 by making temporal ratios explicit in intertemporal choices.
+\subsection{Summary of Studies}
 
-Study 2 extends the results of Study 1 to losses. 
-Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
-For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
-Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
-Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
-Study 2 investigates how framing times as ratios leads to decreased patience in losses.
+In this paper, I describe three studies investigating the explicit temporal ratio effect (ETRE).
+Study 1 shows that the explicit temporal ratios lead to less patience when intertemporal choices are framed as gains.
+Study 2 extends this effect to losses.
+Finally, Study 3 shows that in an explicit shipping scenario people are less patient when temporal ratios are made explicit.
 
-Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
-Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
-Although choosing a shipping speed is similar to the previous intertemporal choices there is a subtle, yet possibly significant difference between the two.
-Shipping decisions can be thought of as paying more now to receive an item sooner whereas the intertemporal choices shown in Study 2 can be thought of as paying something smaller now to avoid paying something larger later.
+%Study 2 extends the results of Study 1 to losses. 
+%Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
+%For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
+%Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
+%Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
+%Study 2 investigates how framing times as ratios leads to decreased patience in losses. 
+%
+%Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
+%Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
+%Although choosing a shipping speed is similar to the previous intertemporal choices there is a subtle, yet possibly significant difference between the two.
+%Shipping decisions can be thought of as \textit{paying more now to receive an item sooner} whereas the intertemporal choices shown in Study 2 can be thought of as \textit{paying something smaller now to avoid paying something larger later}.
+%This difference in choices  
+%
 
 \section{Study 1}
-In study 1 participants chose between smaller sooner and larger later options. 
+Study 1 shows that the explicit temporal ratios lead to less patience when intertemporal choices are framed as gains.
+Participants chose between smaller sooner and larger later options. 
 These options were displayed in one of two frames: implicit temporal ratios and explicit temporal ratios. 
@@
 Additionally only participants who passed an attention check were included in the final analysis.
-These precautions were taken to ensure that participants completed
+These precautions were taken to ensure that participants were paying attention to the study
 
 Participants completed informed consent, then were given a test to ensure they understood the paradigm.
-They then were assigned to one of the two conditions: implicit ratio or explicit ratio. 
-They then were shown the intertemporal choices in Table []. 
-Each intertemporal choice was shown in the following format.
+Next they were assigned to one of the two conditions: implicit ratio or explicit ratio. 
+They then were shown the five intertemporal choices in from the Study 1 section of  Table~\ref{tab:stimuli}. 
+Choices were displayed in a randomized order
 
@@
 \begin{table}[!ht]
-	\caption{Amounts and times of Stimuli for Studies 1-3} \label{tab:stimuli}
+	\caption{Amounts and times of Stimuli for Studies 1-3} 
+	\label{tab:stimuli}
 	\includegraphics[]{Stimuli_For_Study.pdf}
@@
 
-Insert a Qualtrics picture here.
+For the implicit time ratio condition, intertemporal choice A was displayed as follows:
 
-Choices were displayed in a random order to each participant.
+\includegraphics[]{study1_implicit}
 
+Choices B-E were displayed in the identical format.
+
+In the explicit time ratio condition, intertemporal choice A was displayed as follows:
+
+\includegraphics[]{study1_explicit}
+
+Choices B-E were displayed in the identical format.
+
+Further each choice was displayed in a random order to each participant. 
 After completing their choices participants completed demographic information.
@@
 
-The dependent variable in this experiment was the number of times a participant chose the larger later -- e.g. more patient -- reward.
+Following \citeA{Read2005} and \citeA{Hardisty2015}, the dependent variable was the number of times the participant chose sooner reward (option 1 in Table~\ref{tab:stimuli}). 
+As seen the Table~\ref{tab:study1results} (note I did not actually collect data for this study I just put sample data down for illustrative purposes) there is a strong effect of explicit intertemporal ratios.
+Nearly twice as many people choose the sooner option when temporal ratios are explicit.
 
-each participant chose the larger-later reward ?LL?.As Table 2 shows, there was a strong date/delay effect: Twice as many people chose LL in the Date condition than in the Week or Month conditions.
-Differences across conditions were tested formally with Kruskal- Wallis tests.
-An overall test confirmed a signifi- cant difference between the conditions (?2?2?
-= 21?1,
+\begin{table}[!ht]
+	\caption{Proposed Percent of choices for the sooner option given different temporal ratios} 
+	\label{tab:study1results}
+\begin{tabular}{ p{3cm}||p{3cm}|p{3cm}  }
+	\multicolumn{3}{|c|}{Temporal Frame} \\
+	\hline
+	 & Implict Ratio & Explicit Ratio\\
+	\hline
+	A (\%)	 & 30  & 60\\
+	B (\%) 	 & 40  & 65   \\
+	C (\%) 	 & 42  & 70\\
+	D (\%)   & 37  & 68\\
+	E (\%)   & 39  & 66\\
+	Mean (\%) & 38  & 65.8 \\
+	N  	 & 100  & 100\\
+	\hline
+
+\end{tabular}
+\end{table}
+
+Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis tests.
+ Overall the effect of ratio was significant. 
+ To confirm these results, I ran a robust logistic regression which allowed for correlated errors within participants.
+ The dependent variable was choice (0 later option chosen, 1 sooner option chosen) and the independent variable was ratio frame. 
+ This regression confirmed the results of the Kruskal-Wallis test, namely that participants were more impatient when temporal ratios are make explicit.
+
+\section{Study 2}
+
+\section{oldstuff}
 
@@
 
-\begin{table}[!htbp] \centering 
-	\caption{Stimuli Used for Study \_ } \label{tab:stimuli}
-\begin{tabular}{ | l | l | l | l | l | l | l | l | }
-	\hline
-	$SS_{time}$ & $LL_{time}$ & delay & $SS_{time}^{delay}$ & $LL_{time}^{delay}$ & $SS_{amnt}$ & $LL_{amnt}$ & Ratio Change \\ \hline
-	5 & 6 & 5 & 10 & 11 & 105 & 106 & Small  \\ \hline
-	7 & 9 & 17 & 24 & 26 & 122 & 127 & Small  \\ \hline
-	10 & 14 & 14 & 24 & 28 & 144 & 149 & Small  \\ \hline
-	1 & 2 & 5 & 6 & 7 & 170 & 171 & Small  \\ \hline
-	20 & 25 & 22 & 42 & 47 & 108 & 114 & Small  \\ \hline
-	15 & 20 & 25 & 40 & 45 & 140 & 144 & Small  \\ \hline
-	5 & 15 & 5 & 10 & 20 & 105 & 106 & Large \\ \hline
-	7 & 21 & 17 & 24 & 38 & 122 & 127 & Large \\ \hline
-	10 & 22 & 14 & 24 & 36 & 144 & 149 & Large \\ \hline
-	1 & 10 & 5 & 6 & 15 & 170 & 171 & Large \\ \hline
-	20 & 40 & 22 & 42 & 62 & 108 & 114 & Large \\ \hline
-	15 & 75 & 25 & 40 & 100 & 140 & 144 & Large \\ \hline
-\end{tabular}
-\end{table}
+%\begin{table}[!htbp] \centering 
+%	\caption{Stimuli Used for Study \_ } \label{tab:stimuli}
+%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
+%	\hline
+%	$SS_{time}$ & $LL_{time}$ & delay & $SS_{time}^{delay}$ & $LL_{time}^{delay}$ & $SS_{amnt}$ & $LL_{amnt}$ & Ratio Change \\ \hline
+%	5 & 6 & 5 & 10 & 11 & 105 & 106 & Small  \\ \hline
+%	7 & 9 & 17 & 24 & 26 & 122 & 127 & Small  \\ \hline
+%	10 & 14 & 14 & 24 & 28 & 144 & 149 & Small  \\ \hline
+%	1 & 2 & 5 & 6 & 7 & 170 & 171 & Small  \\ \hline
+%	20 & 25 & 22 & 42 & 47 & 108 & 114 & Small  \\ \hline
+%	15 & 20 & 25 & 40 & 45 & 140 & 144 & Small  \\ \hline
+%	5 & 15 & 5 & 10 & 20 & 105 & 106 & Large \\ \hline
+%	7 & 21 & 17 & 24 & 38 & 122 & 127 & Large \\ \hline
+%	10 & 22 & 14 & 24 & 36 & 144 & 149 & Large \\ \hline
+%	1 & 10 & 5 & 6 & 15 & 170 & 171 & Large \\ \hline
+%	20 & 40 & 22 & 42 & 62 & 108 & 114 & Large \\ \hline
+%	15 & 75 & 25 & 40 & 100 & 140 & 144 & Large \\ \hline
+%\end{tabular}
+%\end{table}
 

### 2015-12-15 paper/paperintro.tex
describe losses

@@
 	Other research has shown that people use both differences and ratios of time in intertemporal choices. 
-%	These two results suggest that framing times as ratios alters patience. 
-%	In three experiments I manipulate the temporal presentation from the standard -- 1 day versus 5 days --  to making the ratio explicit -- 1 day or 5 times as long.  
 	Across three experiments I show that making temporal ratios explict leads to differing levels to increase impatience in standard  intertemporal choices framed as gains, intertemporal choices framed as losses, and for an explicit shipping scenario.
 	Taken together these results suggest that framing delays as ratios leads to greater impatience.
-
-%
-%	People use ratios in everyday life -- intertemporal choices are no different, 
-%	recent cognitive models of intertemporal choice show that people calculate ratios of both money and time. 
-%	These models account for diminishing impatience -- when items are moved into the future, people are more patient. 
-%	Our goal is twofold: 1) to show that previous models of intertemporal choice do not properly account for temporal ratios and present a conceptual model of time discounting, and  2) by explicitly manipulating how people use ratios of time, we can alter how people make intertemporal choices.
-%	This leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
-%	Further we show that in a hypothetical shipping scenario that when temporal ratios are explicit, people are more impatient.
 
@@
 Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
-%For instance \citeA{Read2013} found that by making ratios in the 
 However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
 Further although shipping scenarios are a form of intertemporal choice, they differ from how most studies operationalize intertemporal choices.
+
 Across 3 studies I show that by making ratios explicit people are more impatient in intertemporal choices. 
-%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
 This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
@@
 Firms wanting to decrease patience in their consumers may want to frame intertemporal choices as ratios as opposed to delays.
-
-%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
-
-%
-%There are subtle ways in which the presentation of time can alter patience. 
-%For instance people are more impatient when delays are calendar dates as opposed to delays into the future \cite{Read2005}. 
-%Marketers have used temporal frames in shipping scenarios.
-%For example as opposed to stating that you will receive your package in 2 weeks, Amazon sometimes states that you receive your package on January 23rd. 
-%As per \citeA{Read2005} this frame will increase patience.
-
-%Many of our daily choices involve making decisions involving time in some respect, often taking the form of a the tradeoff between receiving something of smaller value now and receiving something of larger value in the future.
-%For instance when a consumer selects a shipping speed when shopping online they are making an intertemporal choice.
-%These decisions require consumers to, perhaps implicitly, calculate how much waiting for the slower shipping speed is worth.
-
-%Firms could, however, prefer that a consumer is impatient when choosing a shipping speed. 
-%If a firm's inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster.
-%It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
-%This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds. 
-%
-%Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
-%Across [] studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
-%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios. 
-%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.
 
@@
 
+Although the majority of intertemporal choice studies involve choices between gains, temporal choices can also take the form of choices between losses \cite{Hardisty2013, Hardisty2015}. 
+Although losses have been shown to be a critical moderator of intertemporal preferences cite{Hardisty2013} few studies explore how loses affect intertemporal choices.
+Further all studies cited so far have only looked at framing effects of intertemporal choices framed as gains. 
+Framing effects which are robust for gains, may have affect losses differently.
+
 This work aims to demonstrate that the fourth research stream -- that patience is affected by how options are framed -- can affect real world phenomenon.
@@
 
-%Study 2 extends the results of Study 1 to losses. 
-%Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
-%For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
-%Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
 %Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
@@
 \section{Study 2}
+
+Study 2 extends the results of Study 1 to losses. 
+Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
+For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
+Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
 

### 2015-12-15 paper/paperintro.tex
Hacking the intro together

@@
 \usepackage{pdflscape}
+\linespread{1.5}
 
@@
 	When selecting a shipping speed on Amazon we are faced with a choice between getting our package sooner but paying more money, or getting our package later and paying less money. 
-	The general phenomenon underlying our expedited shipping is impatience, we want good things immediately and we must be compensated for waiting. 
+	This shipping decision is a subset of a larger class of decisions known as intertemporal choices -- choices involving different points in time. 
 	Previous research on intertemporal framing has shown that displaying times as dates -- January 23rd -- as opposed to delays -- 6 weeks -- increases patience. 
-	Other research has shown that people use both differences and ratios of time in intertemporal choices. 
+	The present research expands this finding. 
 	Across three experiments I show that making temporal ratios explict leads to differing levels to increase impatience in standard  intertemporal choices framed as gains, intertemporal choices framed as losses, and for an explicit shipping scenario.
@@
 
+Intertemporal choices ``involve relative preferences and tradeoffs for costs and benefits that occur over time'' \cite[p. 2]{Zauberman2014}.
+A subset of intertemporal choices are shipping decisions.
 All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible; sometimes leading us choosing expedited shipping, even at considerable cost.
+
 Generally on Amazon we see shipping times as the number of days until you will receive your package, e.g.  ``Receive your package in 2 days''. 
@@
 This reframing of delays -- 2 days -- into dates -- December 17th -- has been shown to increase patience \cite{Read2005}.
-Occasionally Amazon frames shipping times as dates, possibly to induce customers to be more patient and receive their items later.
+Occasionally Amazon frames shipping times as dates, ostensibly to induce more patience from customers.
 
 However a firm may want consumers to be impatient and choose the expedited shipping option. 
-If a firm's inventory is high they may want to nudge consumers to ship their items faster. 
-It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
-This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds.
+If a firm's inventory is high they may want consumers to ship their items faster. 
+The goal of this paper is to introduce how a novel temporal frame -- namely explicit ratios -- which leads to greater degrees of impatience.
+%It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
+%Understanding how to optimally display information leads to novel predictions for how to frame shipping speeds.
 
-How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
-Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
-However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
-Further although shipping scenarios are a form of intertemporal choice, they differ from how most studies operationalize intertemporal choices.
+%How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
+%Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
+%However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
+%Further although shipping scenarios are a form of intertemporal choice, they differ from how most studies operationalize intertemporal choices.
 
@@
 This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
-As opposed to other studies which showed that reframed times as dates increases patience, framing times as ratios decreases patience.
 Firms wanting to decrease patience in their consumers may want to frame intertemporal choices as ratios as opposed to delays.
@@
 
-Intertemporal preferences measured in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}. 
+In marketing and psychology, intertemporal preferences are generally measured by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}. 
 There are four general categories of research on  intertemporal choices (see \citeA{Read2013} for an overview).
 The first focuses on how magnitude of both time and money effects patience \cite{Scholten2006, Thaler1981}
-the current paper belongs to research investigating how patience is altered by option frames. 
 The second stream focuses on how choices relate to real world behavior.
 Individual differences in imputed discount functions predict real world phenomenon such as credit card debt,  mortgage choice, and smoking \cite{MacKillop2011, Meier2009, Johnson2011}. 
-The third stream focuses on how emotional states can alter intertmporal preferences, e.g. being shown freshly baked bread \cite{Li2008}. 
+The third stream focuses on how emotional states can alter intertemporal preferences, e.g. being shown freshly baked bread \cite{Li2008}. 
 The fourth stream investigates how the framing of options affects patience. 
+The current paper is mainly housed within the fourth stream, but attempts to answer how temporal frames alter real world decisions.
 
@@
 Their model -- the Intertemporal Choice Heuristic (ITCH) -- does better at predicting intertemporal choices than classic models, such as hyperbolic or exponential discounting. 
-As seen in Equation~\ref{eq:itch}  the functional form of the ITCH model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting.
+ITCH posits that people weight four attributes -- amount difference, amount ratio, time difference, time ratio --  to varying degrees. 
+If we hold everything else constant and increase the weight on the time ratio component, ITCH predicts that people will be more impatient on identical intertemporal choices. 
+Along with \citeA{Read2013} this suggests that by making temporal ratios more salient will lead to less patience.
 
-[What are they?]
-
- \begin{equation}\label{eq:itch}
- P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
- \end{equation}
-
+%As seen in Equation~\ref{eq:itch}  the functional form of the ITCH model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting. 
+%
+%[What are they?]
+% 
+% \begin{equation}\label{eq:itch}
+% P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
+% \end{equation}
+% 
 Although the majority of intertemporal choice studies involve choices between gains, temporal choices can also take the form of choices between losses \cite{Hardisty2013, Hardisty2015}. 
@@
 
+\subsection{Discussion}
+
+Study 1 demonstrated that explicit temporal ratios lead to greater impatience.
+
 \section{Study 2}
@@
 For instance waiting to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
-Intertemporal choices framed as lossed measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
+Intertemporal choices framed as losses measured in the lab take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
 
-\section{oldstuff}
+\subsection{Method}
+Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
+Participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
+Additionally only participants who passed an attention check were included in the final analysis.
+These precautions were taken to ensure that the sample included qualtiy participants who were paying attention to the study.
 
-Ratios are ubiquitous in the modern world and are paramount in disparate fields from the stock market to baking. 
-When making inferences from ratios, people are biased in predictable ways.
-For instance, Yamagishi showed that ratios with large numerators and denominators are perceived to be riskier than ratios with small numerators and denominators.
+Participants completed informed consent, then were given a test to ensure they understood the paradigm.
+Next they were assigned to one of the two conditions: implicit ratio or explicit ratio. 
+Participants read the instruction
+``Please imagine you face a set of choices
+about paying a bill at a short time from now, or another bill later''
+They then were shown the five intertemporal choices in from the Study 2 section of  Table~\ref{tab:stimuli}. 
+Choices were displayed in a randomized order
+
+For the implicit time ratio condition, intertemporal choice A was displayed as follows:
+
+\includegraphics[]{study2_implicit}
+
+Choices B-E were displayed in the identical format.
+
+In the explicit time ratio condition, intertemporal choice A was displayed as follows:
+
+\includegraphics[]{study2_explicit}
+
+Choices B-E were displayed in the identical format.
+
+Further each choice was displayed in a random order to each participant. 
+After completing their choices participants completed demographic information.
+
+\subsection{Results}
+
+As with Study 1, the dependent variable was the number of times the participant chose sooner reward (option 1 in Table~\ref{tab:stimuli}). 
+As seen the Table~\ref{tab:study2results} (note I did not actually collect data for this study I just put sample data down for illustrative purposes) there is a strong effect of explicit intertemporal ratios.
+Nearly twice as many people choose the sooner option when temporal ratios are explicit.
+
+\begin{table}[!ht]
+	\caption{Proposed Percent of choices in Losses for the sooner option in different temporal ratios} 
+	\label{tab:study2results}
+	\begin{tabular}{ p{3cm}||p{3cm}|p{3cm}  }
+		\multicolumn{3}{|c|}{Temporal Frame} \\
+		\hline
+		& Implict Ratio & Explicit Ratio\\
+		\hline
+		A (\%)	 & 20  & 48\\
+		B (\%) 	 & 25  & 50\\
+		C (\%) 	 & 27  & 57\\
+		D (\%)   & 28  & 54\\
+		E (\%)   & 24  & 58\\
+		Mean (\%) & 38  & 65.8 \\
+		N  	 & 100  & 100\\
+		\hline
+
+	\end{tabular}
+\end{table}
+
+Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis tests.
+Overall the effect of ratio was significant. 
+To confirm these results, I ran a robust logistic regression which allowed for correlated errors within participants.
+The dependent variable was choice (0 later option chosen, 1 sooner option chosen) and the independent variable was ratio frame. 
+This regression confirmed the results of the Kruskal-Wallis test, namely that participants were more impatient when temporal ratios are make explicit.
+
+\subsection{Discussion}
+
+Study 2 replicated the ETRE -- explicit temporal ratios yeild more impatience -- for intertemporal losses.
+
+\section{Study 3}
+Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
+Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
+Although choosing a shipping speed is similar to the intertemporal choices shown in Studies 1 and 2 there is a subtle, yet possibly significant difference between the two.
+Shipping decisions can be thought of as \textit{paying more now to receive an item sooner} whereas the intertemporal choices shown in Study 2 can be thought of as \textit{paying something smaller now to avoid paying something larger later}.
+
+\subsection{Method}
+Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
+Participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
+Additionally only participants who passed an attention check were included in the final analysis.
+These precautions were taken to ensure that the sample included qualtiy participants who were paying attention to the study.
+
+Participants completed informed consent, then were given a test to ensure they understood the paradigm.
+Next they were assigned to one of the two conditions: implicit ratio or explicit ratio. 
+Participants read the instruction
+``Imagine that you have just spent \$75 dollars shopping online.
+None of your purchases are absolutely necessary to have, but you would enjoy having all of them.
+You are about to check out and are presented with two shipping options.''
+They then were shown the five intertemporal choices in from the Study 3 section of  Table~\ref{tab:stimuli}. 
+Choices were displayed in a randomized order
+
+For the implicit time ratio condition, intertemporal choice A was displayed as follows:
+
+\includegraphics[]{study3_implicit}
+
+Choices B-E were displayed in the identical format.
+
+In the explicit time ratio condition, intertemporal choice A was displayed as follows:
+
+\includegraphics[]{study3_explicit}
+
+Choices B-E were displayed in the identical format.
+
+Further each choice was displayed in a random order to each participant. 
+After completing their choices participants completed demographic information.
+
+\subsection{Results}
+
+As with Studies 1 and 2, the dependent variable was the number of times the participant chose sooner reward (option 1 in Table~\ref{tab:stimuli}). 
+As seen the Table~\ref{tab:study3results} (note I did not actually collect data for this study I just put sample data down for illustrative purposes) there is a strong effect of explicit intertemporal ratios.
+Nearly twice as many people choose to expedite their shipping temporal ratios are explicit.
+
+\begin{table}[!ht]
+	\caption{Proposed Percent of choices in Losses for the sooner option in different temporal ratios} 
+	\label{tab:study3results}
+	\begin{tabular}{ p{3cm}||p{3cm}|p{3cm}  }
+		\multicolumn{3}{|c|}{Temporal Frame} \\
+		\hline
+		& Implict Ratio & Explicit Ratio\\
+		\hline
+		A (\%)	 & 20  & 48\\
+		B (\%) 	 & 25  & 50\\
+		C (\%) 	 & 27  & 57\\
+		D (\%)   & 28  & 54\\
+		E (\%)   & 24  & 58\\
+		Mean (\%) & 38  & 65.8 \\
+		N  	 & 100  & 100\\
+		\hline	
+	\end{tabular}
+\end{table}
+
+Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis tests.
+Overall the effect of ratio was significant. 
+To confirm these results, I ran a robust logistic regression which allowed for correlated errors within participants.
+The dependent variable was choice (0 No Rush Shipping Chosen, 1 Expedited Shipping Chosen) and the independent variable was ratio frame. 
+This regression confirmed the results of the Kruskal-Wallis test, namely that participants were more impatient when temporal ratios are make explicit.
+
+\subsection{Discussion}
+
+Study 3 replicated the ETRE -- explicit temporal ratios yield more impatience -- for intertemporal choices explicitly framed as shipping.
+
+\section{General Discussion}
+
+Across three studies we demontrate that the Explicit Temporal Ratio Effect (ETRE) -- making temporal ratios explicit -- yields greater patience. 
+This importance has been understated by previous models of intertemporal choice, and by explicitly expressing temporal ratios people are more impatient.
+
+\section{Limitations and Future Directions}
+
+%\section{oldstuff}
+%
+%
+%Ratios are ubiquitous in the modern world and are paramount in disparate fields from the stock market to baking. 
+%When making inferences from ratios, people are biased in predictable ways.
+%For instance, Yamagishi showed that ratios with large numerators and denominators are perceived to be riskier than ratios with small numerators and denominators.
 Further, Reyna and Brainerd argued that people overweight numerators compared to denominators.
-The biases used when calculating ratios are relevant to intertemporal choices.
-
-Recent models of intertemporal choice have posited that people use ratios to decide between a smaller sooner and a larger later option [ITCH, DRIFT]. 
-An example intertemporal choice is choosing between \$5 in 1 day and \$6 in 5 days. 
-The ITCH model would predict that participants calculate differences between and ratios of both the dollar amounts and the times. 
-If people are biased in calculating ratios for everyday items, it follows that they would be biased in calculating ratios in intertemporal choices.
-
-One of the empirical regularities in intertemporal choice research is that impatience decreases when times are moved away from the present.
-If we took the above example and moved both options back by 20 days we would see more patient responses. 
-Multiple explanations for this phenomenon -- known as decreasing impatience -- have been marshaled.
-
-Zauberman showed that taking into account subjective perceptions of time can account for decreasing impatience. 
-For instance they find that the subjective time perception of 1 year is not different from the subjective time perception of 3 years. 
-Further subjective time perception was shown to follow a psychophysical logarithmic function.
+%The biases used when calculating ratios are relevant to intertemporal choices. 
+%
+%Recent models of intertemporal choice have posited that people use ratios to decide between a smaller sooner and a larger later option [ITCH, DRIFT]. 
+%An example intertemporal choice is choosing between \$5 in 1 day and \$6 in 5 days. 
+%The ITCH model would predict that participants calculate differences between and ratios of both the dollar amounts and the times. 
+%If people are biased in calculating ratios for everyday items, it follows that they would be biased in calculating ratios in intertemporal choices. 
+%
+%One of the empirical regularities in intertemporal choice research is that impatience decreases when times are moved away from the present.
+%If we took the above example and moved both options back by 20 days we would see more patient responses. 
+%Multiple explanations for this phenomenon -- known as decreasing impatience -- have been marshaled. 
+%
+%Zauberman showed that taking into account subjective perceptions of time can account for decreasing impatience. 
+%For instance they find that the subjective time perception of 1 year is not different from the subjective time perception of 3 years. 
+%Further subjective time perception was shown to follow a psychophysical logarithmic function.
 Zauberman, however, embedded this subjective time perception into a hyperbolic discounting function – an alternative based discounting function.
-
-Beyond accounting for diminishing impatience subjective time perception can account for subadditive discounting. 
-Read (2001) found that intertemporal choices are intransitive, specifically “discounting over a delay is greater when the delay is divided into subintervals than when it is left undivided”. 
-However, Zauberman showed that participants perceive the total time horizon as longer when it is divided into subintervals than when it is not divided. 
-Further participants discounted more when the delay was divided into subintervals than when it was undivided.
-
-The focus of this paper is on how people calculate ratios specifically about time. 
-We hypothesize that the ratio change in times alters the extent of decreasing impatience. 
-We begin with a review of the current discounting models and how they account for decreasing impatience. 
-Further we will show that while some models predict that ratio change will have an effect on decreasing impatience, that none of the models predict effects as large as we see. 
-These results introduce a phenomenon whose extent is not captured by the current discounting models. 
-They show that the properties of the numbers given can alter patience via their ratio. 
-Moreover recent models of intertemporal choice do not properly predict the degree of decreasing impatience properly, we show that our modification to their model does a better job of predicting choices.
-
-We then apply our conceptual model to situations where temporal ratios are explicit.
-
-\section{Discounting Models}
-
-When presented with a choice between receiving A) \$10 in 5 days or receiving B) \$12 in 7 days, there are a large number of calculations you could perform to decide which to choose.
+%
+%Beyond accounting for diminishing impatience subjective time perception can account for subadditive discounting. 
+%Read (2001) found that intertemporal choices are intransitive, specifically “discounting over a delay is greater when the delay is divided into subintervals than when it is left undivided”. 
+%However, Zauberman showed that participants perceive the total time horizon as longer when it is divided into subintervals than when it is not divided. 
+%Further participants discounted more when the delay was divided into subintervals than when it was undivided. 
+%
+%The focus of this paper is on how people calculate ratios specifically about time. 
+%We hypothesize that the ratio change in times alters the extent of decreasing impatience. 
+%We begin with a review of the current discounting models and how they account for decreasing impatience. 
+%Further we will show that while some models predict that ratio change will have an effect on decreasing impatience, that none of the models predict effects as large as we see. 
+%These results introduce a phenomenon whose extent is not captured by the current discounting models. 
+%They show that the properties of the numbers given can alter patience via their ratio. 
+%Moreover recent models of intertemporal choice do not properly predict the degree of decreasing impatience properly, we show that our modification to their model does a better job of predicting choices.
+%
+%We then apply our conceptual model to situations where temporal ratios are explicit. 
+%
+%
+%\section{Discounting Models}
+%
+%When presented with a choice between receiving A) \$10 in 5 days or receiving B) \$12 in 7 days, there are a large number of calculations you could perform to decide which to choose.
 Alternative based theories -- such as exponential and hyperbolic discounting -- posit that you calculate a discounted utility for each option.
 Whereas other, attribute based models, posit that people make comparisons attribute based comparisons -- they compare the money and time amounts to one another.
-
-\subsection{Exponential Discounting}
-
-The normative model of intertemporal choice is exponential discounting.
-The exponential discounting function is
-
-\begin{equation}\label{eq:exp}
-	\delta_{exp} = \exp^{-rt}
-\end{equation}
-
-Where $\delta$ is the discount fraction which each monetary amount is multiplied by and $r$ is the exponential discount rate. And the utility of an option is:
-
-\begin{equation}\label{eq:utilalt}
-	U(x, t) = x \delta
-\end{equation}
-
- If the daily exponential discount rate is .01 then the current value of A) Would be $\$10 * \exp^{-.01*5} = 9.5$ and the current value of B) would be $\$12 * \exp^{-.01*7} = 11.19$
-
-From \cite{@Toubia2014} Toubia 2014 we can get the probability of choosing larger later from the following formula
-
-\begin{equation}\label{eq:pll}
-	P_{LL} = \frac{\exp(U_{LL})}{\exp(U_{LL})+ \exp(U_{SS})}
-\end{equation}
-
-Using this formula we get the following probability of choosing larger later $\frac{11.19}{11.19+9.5} = .54$
-
-One of the empirical regularities in intertemporal choice research is that impatience decreases when times are moved away from the present.
+%
+%\subsection{Exponential Discounting}
+%
+%The normative model of intertemporal choice is exponential discounting.
+%The exponential discounting function is  
+%
+%\begin{equation}\label{eq:exp}
+%	\delta_{exp} = \exp^{-rt}
+%\end{equation}
+%
+%
+%Where $\delta$ is the discount fraction which each monetary amount is multiplied by and $r$ is the exponential discount rate. And the utility of an option is:
+%
+%\begin{equation}\label{eq:utilalt}
+%	U(x, t) = x \delta
+%\end{equation}
+%
+%
+%
+% If the daily exponential discount rate is .01 then the current value of A) Would be $\$10 * \exp^{-.01*5} = 9.5$ and the current value of B) would be $\$12 * \exp^{-.01*7} = 11.19$
+%
+%From \cite{@Toubia2014} Toubia 2014 we can get the probability of choosing larger later from the following formula
+%
+%\begin{equation}\label{eq:pll}
+%	P_{LL} = \frac{\exp(U_{LL})}{\exp(U_{LL})+ \exp(U_{SS})}
+%\end{equation}
+%
+%
+%Using this formula we get the following probability of choosing larger later $\frac{11.19}{11.19+9.5} = .54$
+%
+%One of the empirical regularities in intertemporal choice research is that impatience decreases when times are moved away from the present.
 The exponential model, however, does not account for this phenomenon.
@@
 Plugging into equation \ref{eq:pll} we get $.54$, the same answer we saw before.
-
-\subsection{Hyperbolic}
-
-Hyperbolic discounting has been shown to account for decreasing impatience
-
-\begin{equation}\label{eq:hyp}
-\delta_{hyp} = \frac{1}{1+kt}
-\end{equation}
-
-Where $k$ is the hyperbolic discount rate, and the utility of the option is found by equation \ref{eq:utilalt}.
-
-Assuming daily $k = .01$, for choice A) \ref{eq:hyp} we get $\delta_A = \frac{1}{1+.01*5} = .952$ meaning the utility is $9.52$ and for B) the utility is $11.21$ which yields a larger later probability of $.541$.
+%
+%\subsection{Hyperbolic}
+%
+%Hyperbolic discounting has been shown to account for decreasing impatience
+%
+%\begin{equation}\label{eq:hyp}
+%\delta_{hyp} = \frac{1}{1+kt}
+%\end{equation}
+%
+%Where $k$ is the hyperbolic discount rate, and the utility of the option is found by equation \ref{eq:utilalt}.
+%
+%Assuming daily $k = .01$, for choice A) \ref{eq:hyp} we get $\delta_A = \frac{1}{1+.01*5} = .952$ meaning the utility is $9.52$ and for B) the utility is $11.21$ which yields a larger later probability of $.541$.
 When we look at the larger later probabilities for C and D we get, 4.88 and  5.79 yielding a larger later value of $.543$.
 There's an increase in the likelihood of choosing the larger later option when both options are moved into the future, this is known as diminishing impatience.
-
-The hyperbolic and exponential models both predict that people make alternative based transitions; however recent models have shown that people make attribute based transitions
-
-\subsubsection{Subjective Time Perception}
-
-Zauberman showed that peoples subjective perception of time can account for hyperbolic discounting.
-
-\subsection{ITCH}
-\label{ITCH}
-
-Marzilli Ericson and colleagues showed that fully attribute based models can predict out of sample choices better than alternative based models. 
-According to ITCH the probability of choosing larger later is
-
-%\begin{equation}\label{eq:itch}
-%	P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
-%\end{equation}
-
-In the paper they give the example parameters for each of the coefficients of 0, .1, .1.
+%
+%The hyperbolic and exponential models both predict that people make alternative based transitions; however recent models have shown that people make attribute based transitions
+%
+%
+%
+%\subsubsection{Subjective Time Perception}
+%
+%Zauberman showed that peoples subjective perception of time can account for hyperbolic discounting. 
+%
+%\subsection{ITCH}
+%\label{ITCH}
+%
+%Marzilli Ericson and colleagues showed that fully attribute based models can predict out of sample choices better than alternative based models. 
+%According to ITCH the probability of choosing larger later is
+%
+%%\begin{equation}\label{eq:itch}
+%%	P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
+%%\end{equation}
+%
+%In the paper they give the example parameters for each of the coefficients of 0, .1, .1.
 -.1, -.1.
 Using these coefficients in equation \ref{eq:itch} for A compared to B the probability of choosing larger later would be $.496$, whereas for C and D the probability of choosing larger later would be $.504$
-
-\subsection{Discounting By Intervals}
-
-\begin{equation}\label{eq:int}
-	D(t_S, t_L) = \left(1 + \alpha \left(\frac{w(t_L) - w(t_S)}{\vartheta}\right)^{\vartheta}\right)^{-\beta/\alpha}
-\end{equation}
-
-\subsection{Trade-off model}
-
-% %\begin{equation}
-% %Q(w(t_L) - w(t_S)) = 
-% %\begin{cases}
-% %v(x_L) - v(x_S), & \text{if}\ x_L > x_S > 0 \\
-% %v(x_S) - v(x_L), & \text{if}\ 0 > x_S > x_L
-% %\end{cases}
-% %\end{equation}
-
-\subsection{Summary}
-Taken together these models show that some models predict that a large ratio change will have more decreasing impatience than a small ratio change; however we posit that the degree of  decreasing impatience will be larger when the ratio change is small compared to when it is large.
-Take the following two choices for example $SS_{sr}$ = \$10 in 5 days and $LL_{sr}$ = \$12 in 6 days and $SS_{lr}$ = \$10 in 5 days and $LL_{lr}$ = \$12 in 10 days.
+%
+%
+%\subsection{Discounting By Intervals}
+%
+%\begin{equation}\label{eq:int}
+%	D(t_S, t_L) = \left(1 + \alpha \left(\frac{w(t_L) - w(t_S)}{\vartheta}\right)^{\vartheta}\right)^{-\beta/\alpha}
+%\end{equation}
+%
+%
+%\subsection{Trade-off model}
+%
+%% %\begin{equation}
+%% %Q(w(t_L) - w(t_S)) = 
+%% %\begin{cases}
+%% %v(x_L) - v(x_S), & \text{if}\ x_L > x_S > 0 \\
+%% %v(x_S) - v(x_L), & \text{if}\ 0 > x_S > x_L
+%% %\end{cases}
+%% %\end{equation}
+%
+%\subsection{Summary}
+%Taken together these models show that some models predict that a large ratio change will have more decreasing impatience than a small ratio change; however we posit that the degree of  decreasing impatience will be larger when the ratio change is small compared to when it is large.
+%Take the following two choices for example $SS_{sr}$ = \$10 in 5 days and $LL_{sr}$ = \$12 in 6 days and $SS_{lr}$ = \$10 in 5 days and $LL_{lr}$ = \$12 in 10 days.
 Where $SS_{sr}$ stands for the smaller sooner option for the small time ratio.
-If we add 5 days to both items they would now be $SS_{sr}^{del}$ = \$10 in 10 days and $LL_{sr}^{del}$ = \$12 in 11 days and $SS_{lr}^{del}$ = \$10 in 10 days and $LL_{lr}^{del}$ = \$12 in 15 days.
-
-[Insert a table about the degree of decreasing impatience for each model here]
-
-\section{Time Ratio Model (TRaM)}
-
-We propose a new model of ITC which can account for the degree of decreasing impatience based on the ratios of time. 
-We begin with a verbal description of our model
-
-\subsection{An example of TRaM}
-
-When presented with a choice between \$5 in 1 day and \$10 in 5 days. 
-A TRM decision maker would look at the two delays and calcluate their ratio. 
-Then they will use a rule of thumb to say how much relative compensation they need to account for that ratio.
-Below is an example of a TRM decision maker would make an intertemporal choice:
-
-The time is 5 times as long therefore I will need about 3 times as much money.
-
-If people use this heuristic then the ratio between times becomes important. 
-If we move both options up by five days then the ratio between times becomes $10/6$ = 1.66. A participant using the TRM heuristic would say that 
-The larger later time is about 1.5 times as long as the smaller sooner time therefore I need about 1.25 times as much money.
-
-Further TRaM predicts that the salience of the temporal ratio means that it is more likely to be used. 
-For instance if the temporal ratio explicit, participants reliance on it would increase.
-
-\section{Combining ITCH, DRIFT, and TRaM }
-
-Read, Frederick, and Scholten (2013) show that when making intertemporal choices how the monetary amounts are framed alters patience. 
-They propose a Difference, Ratio, Interest, Finance, and Time (DRIFT) model of intertemporal choice. 
-For instance when an intertemporal choice is framed as \$100 now OR \$110 in 1 year people will rely on differences in monetary amounts.
-However if the choice is framed as \$100 now or an extra 10 \% after 1 year, people will rely more on ratios. 
-The DRIFT model of intertemporal choice shows that how monetary amounts are framed can alter levels of patience.
-
-Relatedly Read, Frederick, Orsel, and Rahman showed that how the times intertemporal choice are displayed can affect patience.
+%If we add 5 days to both items they would now be $SS_{sr}^{del}$ = \$10 in 10 days and $LL_{sr}^{del}$ = \$12 in 11 days and $SS_{lr}^{del}$ = \$10 in 10 days and $LL_{lr}^{del}$ = \$12 in 15 days. 
+%
+%[Insert a table about the degree of decreasing impatience for each model here]
+%
+%\section{Time Ratio Model (TRaM)}
+%
+%We propose a new model of ITC which can account for the degree of decreasing impatience based on the ratios of time. 
+%We begin with a verbal description of our model
+%
+%\subsection{An example of TRaM}
+%
+%When presented with a choice between \$5 in 1 day and \$10 in 5 days. 
+%A TRM decision maker would look at the two delays and calcluate their ratio. 
+%Then they will use a rule of thumb to say how much relative compensation they need to account for that ratio.
+%Below is an example of a TRM decision maker would make an intertemporal choice:
+%
+%The time is 5 times as long therefore I will need about 3 times as much money. 
+%
+%If people use this heuristic then the ratio between times becomes important. 
+%If we move both options up by five days then the ratio between times becomes $10/6$ = 1.66. A participant using the TRM heuristic would say that 
+%The larger later time is about 1.5 times as long as the smaller sooner time therefore I need about 1.25 times as much money. 
+%
+%Further TRaM predicts that the salience of the temporal ratio means that it is more likely to be used. 
+%For instance if the temporal ratio explicit, participants reliance on it would increase.
+%
+%
+%
+%\section{Combining ITCH, DRIFT, and TRaM }
+%
+%Read, Frederick, and Scholten (2013) show that when making intertemporal choices how the monetary amounts are framed alters patience. 
+%They propose a Difference, Ratio, Interest, Finance, and Time (DRIFT) model of intertemporal choice. 
+%For instance when an intertemporal choice is framed as \$100 now OR \$110 in 1 year people will rely on differences in monetary amounts.
+%However if the choice is framed as \$100 now or an extra 10 \% after 1 year, people will rely more on ratios. 
+%The DRIFT model of intertemporal choice shows that how monetary amounts are framed can alter levels of patience.
+%
+%Relatedly Read, Frederick, Orsel, and Rahman showed that how the times intertemporal choice are displayed can affect patience.
 Specifically Read and colleagues found that when times are described as dates (e.g.
@@
 Temporal framing can alter patience.
-
-Given that temporal framing can alter patience and that people rely on ratiosin intertemporal choice, we propose that making ratios between times explicit can alter patience. 
-For instance, an intertemporal choice between receiving \$10 in 2 days and \$11 in 12 days could be re-framed as \$10 in 2 days and \$11 in 6 times as long. 
-We proposed that this temporal framing leads to a greater weight being placed upon the time ratio term $\beta_{tR}$ and a lower weight on the time difference term $\beta_{tA}$ in equation \ref{eq:itch} .
+%
+%Given that temporal framing can alter patience and that people rely on ratiosin intertemporal choice, we propose that making ratios between times explicit can alter patience. 
+%For instance, an intertemporal choice between receiving \$10 in 2 days and \$11 in 12 days could be re-framed as \$10 in 2 days and \$11 in 6 times as long. 
+%We proposed that this temporal framing leads to a greater weight being placed upon the time ratio term $\beta_{tR}$ and a lower weight on the time difference term $\beta_{tA}$ in equation \ref{eq:itch} .
 Taking the example and using the parameters in section \ref{ITCH} the probability of choosing larger later is $.26$.
@@
 $\beta_{tA} = -.1$, the probability of choosing larger later is now $.0065$.
-
-Holding everything else constant by making the temporal ratio more salient, e.g.
+%
+%Holding everything else constant by making the temporal ratio more salient, e.g.
 increasing the absolute value of $\beta_{tR}$,  monotonically decreases the probability of choosing larger later.
 We propose that making temporal ratios more salient increases their decision weight, leading to more impatient choices
-
-{\large [This CANNOT account for the change in the decision weight on temporal differences by making temporal ratios salient, which I am unsure of how to work out]. 
-}
-
-The TRaM conceptual model of intertemporal choice suggests that by making temporal ratios explicit that people's reliance on them will be higher.
-When people's reliance on temporal ratios is higher, the weight in which temporal ratios receive will be higher. 
-This weight has predictable consequences on how impatient participants are, which suggests ways in which firms can alter patience based on how times are presented.
-
-\section{Study 1}
-
-In study 1 we will test this prediction of the degree of decreasing impatience depending on the change in ratio.
-We predict that when ratio changes are large, participants will show more diminishing impatience than when ratio changes are small.
-
-\subsection{Methods}
-Participants were in one of two conditions. 
-Six intertemporal choices were consistent across conditions and were used to calculate the participants discount fraction.
-The other two intertemporal choices varied across conditions, these choices varied in the extent to which the temporal ratio changed.
-We varied the ratio with either a small or large time ratio which was moved into the future by 5 days.
-
-In condition 1 participants the two focal questions were:  
-\$10 in 5 days or \$11 in 6 days
-and 
-\$10 in 10 days or \$11 in 11 days
-
-In condition 2 participants the two focal questions were: 
-\$10 in 5 days or \$11 in 15 days
-and 
-\$10 in 10 days or \$11 in 20 days
-
-In condition 1 the change in temporal ratio is small (1.2 to 1.1), whereas in condition 2 the change in temporal ratio is large (3 to 2).
-
-\subsubsection{Participants}
-
-We recruited [] Rutgers university undergraduates to complete the study for course credit.
-
-\subsection{Results}
-
-We calculated the daily discount fraction for each of the four focal items while controlling for the participants discount fraction from the other 6 items. 
-The degree of decreasing impatience was calculated by taking the difference between the daily discount fraction from the two focal items. 
-We ran a multilevel linear regression predicting the degree of decreasing impatience with condition and discount fraction of the other items as predictors.
-
-The regression showed that the degree of decreasing impatience was predicted by condition.
-
-\subsection{Discussion}
-
-Study 1 shows that the use of temporal ratios is more important than other cognitive models predicted. 
-If temporal ratios are important, by explicitly manipulating them, we should be able to alter levels of patience
-
-\section{Study 2}
-
-In study 2 we manipulate how easy it is to compare temporal ratios. 
-From ITCH and TRaM we hypothesize that participants who are shown explicit temporal ratios will make more impatient decisions
-
-\subsection{Methods}
-
-Each respondent answered 27 questions framed as intertemporal choices.
-We varied the time to the smaller sooner option ($SS_{time} =$ 5, 10, or 50 days), 
-the ratio between the smaller sooner and larger later times ($SS_{time} / LL_{time}  = $ 1.2, 2.2, or 3.2 times as long), 
- the difference between the smaller sooner and larger later amount ($LL_{amnt} - SS_{amnt} =$ \$5, \$10, \$30). All smaller sooner amounts were \$47 [Or we could pull from a uniform distribution?].
-
-Each respondent was assigned to one of two frames: implicit temporal ratios and explicit temporal ratios. An example of both is given below
-
-\textbf{Implicit Temporal Ratio: } You can receive \$47 in 5 days or \$54 in 6 days
-
-\textbf{Explicit Temporal Ratio: } You can receive \$47 in 5 days or \$54 in 1.2 times as long
-
-\subsection{Results}
-
-We perform a multilevel logistic regression predicting choice with varying intercepts for each participant, the three within subject factors, and the between subject factor of temporal ratio frame.
-
-The results of the logistic regression are shown in table [].
+%
+%{\large [This CANNOT account for the change in the decision weight on temporal differences by making temporal ratios salient, which I am unsure of how to work out]. 
+%}
+%
+%The TRaM conceptual model of intertemporal choice suggests that by making temporal ratios explicit that people's reliance on them will be higher.
+%When people's reliance on temporal ratios is higher, the weight in which temporal ratios receive will be higher. 
+%This weight has predictable consequences on how impatient participants are, which suggests ways in which firms can alter patience based on how times are presented. 
+%
+%
+%
+%
+%\section{Study 1}
+%
+%In study 1 we will test this prediction of the degree of decreasing impatience depending on the change in ratio.
+%We predict that when ratio changes are large, participants will show more diminishing impatience than when ratio changes are small. 
+%
+%\subsection{Methods}
+%Participants were in one of two conditions. 
+%Six intertemporal choices were consistent across conditions and were used to calculate the participants discount fraction.
+%The other two intertemporal choices varied across conditions, these choices varied in the extent to which the temporal ratio changed.
+%We varied the ratio with either a small or large time ratio which was moved into the future by 5 days. 
+%
+%In condition 1 participants the two focal questions were:  
+%\$10 in 5 days or \$11 in 6 days
+%and 
+%\$10 in 10 days or \$11 in 11 days
+%
+%In condition 2 participants the two focal questions were: 
+%\$10 in 5 days or \$11 in 15 days
+%and 
+%\$10 in 10 days or \$11 in 20 days
+%
+%In condition 1 the change in temporal ratio is small (1.2 to 1.1), whereas in condition 2 the change in temporal ratio is large (3 to 2). 
+%
+%
+%\subsubsection{Participants}
+%
+%We recruited [] Rutgers university undergraduates to complete the study for course credit. 
+%
+%\subsection{Results}
+% 
+%We calculated the daily discount fraction for each of the four focal items while controlling for the participants discount fraction from the other 6 items. 
+%The degree of decreasing impatience was calculated by taking the difference between the daily discount fraction from the two focal items. 
+%We ran a multilevel linear regression predicting the degree of decreasing impatience with condition and discount fraction of the other items as predictors. 
+%
+%The regression showed that the degree of decreasing impatience was predicted by condition.  
+%
+%\subsection{Discussion}
+%
+%Study 1 shows that the use of temporal ratios is more important than other cognitive models predicted. 
+%If temporal ratios are important, by explicitly manipulating them, we should be able to alter levels of patience
+%
+%\section{Study 2}
+%
+%In study 2 we manipulate how easy it is to compare temporal ratios. 
+%From ITCH and TRaM we hypothesize that participants who are shown explicit temporal ratios will make more impatient decisions 
+%
+%\subsection{Methods}
+%
+%Each respondent answered 27 questions framed as intertemporal choices.
+%We varied the time to the smaller sooner option ($SS_{time} =$ 5, 10, or 50 days), 
+%the ratio between the smaller sooner and larger later times ($SS_{time} / LL_{time}  = $ 1.2, 2.2, or 3.2 times as long), 
+% the difference between the smaller sooner and larger later amount ($LL_{amnt} - SS_{amnt} =$ \$5, \$10, \$30). All smaller sooner amounts were \$47 [Or we could pull from a uniform distribution?].
+%
+%Each respondent was assigned to one of two frames: implicit temporal ratios and explicit temporal ratios. An example of both is given below
+%
+%\textbf{Implicit Temporal Ratio: } You can receive \$47 in 5 days or \$54 in 6 days
+%
+%\textbf{Explicit Temporal Ratio: } You can receive \$47 in 5 days or \$54 in 1.2 times as long
+%
+%\subsection{Results}
+%
+%We perform a multilevel logistic regression predicting choice with varying intercepts for each participant, the three within subject factors, and the between subject factor of temporal ratio frame. 
+%
+%The results of the logistic regression are shown in table [].
 We see that participants in the explicit temporal ratio are more impatient frame than those in the implicit temporal ratio frame.
@@
 we see this effect only when the ratio is large but not when it is small?]
-
-\section{Study 3}
-
-While study 2 showed that when time ratios are explicit, people are more impatient, it was framed as gains not losses.
-Most shipping situations involve a cost to the consumer. 
-In Study 3 we replicate study 2 but with losses.
-
-\subsection{Methods}
-
-The methods for study 3 are identical to study 2 except that all monetary amounts are framed as losses and not gains.
-
-\subsection{Results}
-
-We replicate the findings of study 2 when framed as losses.
-
-\section{Study 4}
-
-Studies 1 and 2 showed that explicit time ratios make people more impatient for both gains and losses, but do not specifically deal with a shipping scenario.
+%
+%
+%\section{Study 3}
+%
+%While study 2 showed that when time ratios are explicit, people are more impatient, it was framed as gains not losses.
+%Most shipping situations involve a cost to the consumer. 
+%In Study 3 we replicate study 2 but with losses.
+%
+%\subsection{Methods}
+%
+%The methods for study 3 are identical to study 2 except that all monetary amounts are framed as losses and not gains. 
+%
+%\subsection{Results}
+%
+%We replicate the findings of study 2 when framed as losses. 
+%
+%\section{Study 4}
+%
+%Studies 1 and 2 showed that explicit time ratios make people more impatient for both gains and losses, but do not specifically deal with a shipping scenario.
 Study 4 uses a hypothetical shipping scenario.
-
-\subsection{Methods}
-
-Participants were told to imagine that they had just decided to buy \$75 worth of merchandise on a popular on-line vendor. 
-[Should this strictly be between subjects?]
-
-They then were presented with two shipping options in one of the two frames.
-
-\textbf{Implicit Temporal Ratio}
-
-Option 1: Receive your order in two days for \$7
-
-Option 2: Receive your order in 10 days for \$3
-
-\textbf{Explicit Temporal Ratio}
-
-Option 1: Receive your order in two days for \$7
-
-Option 2: Receive your order in 5 times longer than option 1 for \$3
-
-\subsection{Results}
-
-We perform a logistic regression with shipping choice as the dependent measure and explicit vs implicit temporal ratio as the independent variable. 
-We find that participants are more likely to delay their shipping in the implicit temporal ratio condition.
-
-\subsection{Discussion}
-
-Study 4 shows that temporal ratios are an important driver of real world shipping choices. 
-By making temporal ratios explicit, people are more likely to be patient.
-
-\section{Study}
-
-\subsection{Methods}
-
-\subsubsection{Participants}
-
-Ninety three Rutgers students took part in this study as a part of their course requirements.
-
-\subsubsection{Design}
-
-Participants were randomly assigned to one of two conditions -- small ratios between times and large ratios between times.  
-As seen in Table~\ref{tab:stimuli}, 24 unique intertemporal choices were created. 
-The first column shows the first two times, the third column shows the delay added to both of the first two columns which yeild columns 4 and 5. 
-The $SS_{amnt}$ and $LL_{amnt}$ columns were chosen using intuition to generate half of the participants choosing the smaller sooner option and half choosing the larger later option. 
-The last column indicates the condition which participants saw, e.g. small ratio versus large ratio change
-
-%\begin{table}[!htbp] \centering 
-%	\caption{Stimuli Used for Study \_ } \label{tab:stimuli}
-%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
-%	\hline
-%	$SS_{time}$ & $LL_{time}$ & delay & $SS_{time}^{delay}$ & $LL_{time}^{delay}$ & $SS_{amnt}$ & $LL_{amnt}$ & Ratio Change \\ \hline
-%	5 & 6 & 5 & 10 & 11 & 105 & 106 & Small  \\ \hline
-%	7 & 9 & 17 & 24 & 26 & 122 & 127 & Small  \\ \hline
-%	10 & 14 & 14 & 24 & 28 & 144 & 149 & Small  \\ \hline
-%	1 & 2 & 5 & 6 & 7 & 170 & 171 & Small  \\ \hline
-%	20 & 25 & 22 & 42 & 47 & 108 & 114 & Small  \\ \hline
-%	15 & 20 & 25 & 40 & 45 & 140 & 144 & Small  \\ \hline
-%	5 & 15 & 5 & 10 & 20 & 105 & 106 & Large \\ \hline
-%	7 & 21 & 17 & 24 & 38 & 122 & 127 & Large \\ \hline
-%	10 & 22 & 14 & 24 & 36 & 144 & 149 & Large \\ \hline
-%	1 & 10 & 5 & 6 & 15 & 170 & 171 & Large \\ \hline
-%	20 & 40 & 22 & 42 & 62 & 108 & 114 & Large \\ \hline
-%	15 & 75 & 25 & 40 & 100 & 140 & 144 & Large \\ \hline
-%\end{tabular}
-%\end{table}
-
-\subsubsection{Procedure}
-
-Participants initially granted informed consent. 
-Then were randomly selected to be in one of the two conditions Small or Large ratio change. 
-Participants then made ``unlikeness'' judgments between each set of times in their condition. 
-Furhter, non-delayed and delayed times were blocked together the order in which the questions were asked and the order in which the blocks were presented was randomized. 
-The blocking of items was to minimize the likelihood that participants saw items in the same row of Table~\ref{tab:stimuli} consecutively. 
-For instance a participant in the small ratio change condition answered how unlike each  the 6 non-delayed times were then answered the how unlike eachother each of the 6 delayed times were.
-
-Participants then completed an unrelated task about buying cars, which will not be reported further.
-
-Next particpants answered each of the intertemporal choices in theri condition. 
-The intertemporal choices were blocked similarly to the unlikeness judgments. 
-Next participants were asked the extent to which they used ratios and the extent to which they used differences when making intertemporal choices. 
-Finally participants answered demographics questions.
-
-\subsection{Results}
-
-\section{General Discussion}
-
-Across four studies we show temporal ratios are an important driver of temporal choice. 
-This importance has been understated by previous models of intertemporal choice, and by explicitly expressing temporal ratios people are more impatient.
-
-\section{Limitations and Future Directions}
+%
+%\subsection{Methods}
+%
+%Participants were told to imagine that they had just decided to buy \$75 worth of merchandise on a popular on-line vendor. 
+%[Should this strictly be between subjects?]
+%
+%They then were presented with two shipping options in one of the two frames.
+%
+%\textbf{Implicit Temporal Ratio}
+%
+%Option 1: Receive your order in two days for \$7
+%
+%Option 2: Receive your order in 10 days for \$3
+% 
+% 
+%\textbf{Explicit Temporal Ratio}
+%
+%Option 1: Receive your order in two days for \$7
+%
+%Option 2: Receive your order in 5 times longer than option 1 for \$3
+%
+%\subsection{Results}
+%
+%We perform a logistic regression with shipping choice as the dependent measure and explicit vs implicit temporal ratio as the independent variable. 
+%We find that participants are more likely to delay their shipping in the implicit temporal ratio condition. 
+%
+%\subsection{Discussion}
+%
+%Study 4 shows that temporal ratios are an important driver of real world shipping choices. 
+%By making temporal ratios explicit, people are more likely to be patient. 
+%
+%\section{Study}
+%
+%\subsection{Methods}
+%
+%\subsubsection{Participants}
+%
+%Ninety three Rutgers students took part in this study as a part of their course requirements. 
+%
+%\subsubsection{Design}
+%
+%Participants were randomly assigned to one of two conditions -- small ratios between times and large ratios between times.  
+%As seen in Table~\ref{tab:stimuli}, 24 unique intertemporal choices were created. 
+%The first column shows the first two times, the third column shows the delay added to both of the first two columns which yeild columns 4 and 5. 
+%The $SS_{amnt}$ and $LL_{amnt}$ columns were chosen using intuition to generate half of the participants choosing the smaller sooner option and half choosing the larger later option. 
+%The last column indicates the condition which participants saw, e.g. small ratio versus large ratio change
+%
+%%\begin{table}[!htbp] \centering 
+%%	\caption{Stimuli Used for Study \_ } \label{tab:stimuli}
+%%\begin{tabular}{ | l | l | l | l | l | l | l | l | }
+%%	\hline
+%%	$SS_{time}$ & $LL_{time}$ & delay & $SS_{time}^{delay}$ & $LL_{time}^{delay}$ & $SS_{amnt}$ & $LL_{amnt}$ & Ratio Change \\ \hline
+%%	5 & 6 & 5 & 10 & 11 & 105 & 106 & Small  \\ \hline
+%%	7 & 9 & 17 & 24 & 26 & 122 & 127 & Small  \\ \hline
+%%	10 & 14 & 14 & 24 & 28 & 144 & 149 & Small  \\ \hline
+%%	1 & 2 & 5 & 6 & 7 & 170 & 171 & Small  \\ \hline
+%%	20 & 25 & 22 & 42 & 47 & 108 & 114 & Small  \\ \hline
+%%	15 & 20 & 25 & 40 & 45 & 140 & 144 & Small  \\ \hline
+%%	5 & 15 & 5 & 10 & 20 & 105 & 106 & Large \\ \hline
+%%	7 & 21 & 17 & 24 & 38 & 122 & 127 & Large \\ \hline
+%%	10 & 22 & 14 & 24 & 36 & 144 & 149 & Large \\ \hline
+%%	1 & 10 & 5 & 6 & 15 & 170 & 171 & Large \\ \hline
+%%	20 & 40 & 22 & 42 & 62 & 108 & 114 & Large \\ \hline
+%%	15 & 75 & 25 & 40 & 100 & 140 & 144 & Large \\ \hline
+%%\end{tabular}
+%%\end{table}
+%
+%\subsubsection{Procedure}
+%
+%Participants initially granted informed consent. 
+%Then were randomly selected to be in one of the two conditions Small or Large ratio change. 
+%Participants then made ``unlikeness'' judgments between each set of times in their condition. 
+%Furhter, non-delayed and delayed times were blocked together the order in which the questions were asked and the order in which the blocks were presented was randomized. 
+%The blocking of items was to minimize the likelihood that participants saw items in the same row of Table~\ref{tab:stimuli} consecutively. 
+%For instance a participant in the small ratio change condition answered how unlike each  the 6 non-delayed times were then answered the how unlike eachother each of the 6 delayed times were. 
+%
+%Participants then completed an unrelated task about buying cars, which will not be reported further.
+%
+%Next particpants answered each of the intertemporal choices in theri condition. 
+%The intertemporal choices were blocked similarly to the unlikeness judgments. 
+%Next participants were asked the extent to which they used ratios and the extent to which they used differences when making intertemporal choices. 
+%Finally participants answered demographics questions. 
+%
+%\subsection{Results}
+%
+%\section{General Discussion}
+%
+%Across four studies we show temporal ratios are an important driver of temporal choice. 
+%This importance has been understated by previous models of intertemporal choice, and by explicitly expressing temporal ratios people are more impatient.
 

### 2015-12-15 paper/paperintro.tex
Tightening up...

@@
 This reframing of delays -- 2 days -- into dates -- December 17th -- has been shown to increase patience \cite{Read2005}.
-Occasionally Amazon frames shipping times as dates, ostensibly to induce more patience from customers.
+Occasionally Amazon uses the date frame, ostensibly to induce more patience from customers.
 
@@
 
-Across 3 studies I show that by making ratios explicit people are more impatient in intertemporal choices. 
+Across 3 studies I show that by making implicit temporal ratios explicit people are more impatient in intertemporal choices. 
 This result holds for intertemporal choices framed as gains, losses, and in an explicit shipping scenario. 
@@
 
-In marketing and psychology, intertemporal preferences are generally measured by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}. 
 There are four general categories of research on  intertemporal choices (see \citeA{Read2013} for an overview).
-The first focuses on how magnitude of both time and money effects patience \cite{Scholten2006, Thaler1981}
+The first stream focuses on how magnitude of both time and money effects patience \cite{Scholten2006, Thaler1981}
 The second stream focuses on how choices relate to real world behavior.
@@
 The fourth stream investigates how the framing of options affects patience. 
-The current paper is mainly housed within the fourth stream, but attempts to answer how temporal frames alter real world decisions.
+The current paper is mainly housed within the fourth stream, but also attempts to answer how temporal frames alter real world decisions.
 
@@
 
-In other work on intertemporal choices, \citeA{MarzilliEricson2015} show that in intertemporal choices framed as ``receiving \$100 now'' and ``\$110 in 6 months'', people rely on both differences and ratios of the  attributes of intertemporal choices -- monetary amounts and times.
+In work on the cognitive processes underlying intertemporal choices, \citeA{MarzilliEricson2015} show that in intertemporal choices framed as ``receiving \$100 now'' and ``\$110 in 6 months'', people rely on both differences and ratios of the  attributes of intertemporal choices -- monetary amounts and times.
 Their model -- the Intertemporal Choice Heuristic (ITCH) -- does better at predicting intertemporal choices than classic models, such as hyperbolic or exponential discounting. 
-ITCH posits that people weight four attributes -- amount difference, amount ratio, time difference, time ratio --  to varying degrees. 
-If we hold everything else constant and increase the weight on the time ratio component, ITCH predicts that people will be more impatient on identical intertemporal choices. 
-Along with \citeA{Read2013} this suggests that by making temporal ratios more salient will lead to less patience.
+ITCH posits that people weight four attributes -- amount difference, amount ratio, time difference, time ratio --  to varying degrees when making intertemporal choices. 
+Holding everything else constant and increasing the weight placed on the time ratio component, ITCH predicts that people will be more impatient on identical intertemporal choices.
 
-%As seen in Equation~\ref{eq:itch}  the functional form of the ITCH model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting. 
-%
-%[What are they?]
-% 
-% \begin{equation}\label{eq:itch}
-% P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
-% \end{equation}
-% 
-Although the majority of intertemporal choice studies involve choices between gains, temporal choices can also take the form of choices between losses \cite{Hardisty2013, Hardisty2015}. 
-Although losses have been shown to be a critical moderator of intertemporal preferences cite{Hardisty2013} few studies explore how loses affect intertemporal choices.
-Further all studies cited so far have only looked at framing effects of intertemporal choices framed as gains. 
-Framing effects which are robust for gains, may have affect losses differently.
+Taken together, this suggests that by making temporal ratios more salient will lead to less patience.
+I propose a new framing effect in intertemporal choice: the Explicit Temporal Ratio Effect (ETRE). 
+The ETRE posits that by making temporal ratios salient people will be more impatient. 
+Specifically, the ETRE focuses on how framing times as ratios can decrease patience.
+When delays are sufficiently small, and when temporal ratios are made explicit, our attention will be focused on how much \textit{relatively longer} we will wait. 
+For instance reframing a choice between ``receiving \$100  in 1 day'' and ``\$110 in \textit{10 days}'' to being a choice between ``receiving \$100  in 1 day'' and ``\$110 in 10 \textit{times as long}'', will focus participants on how much longer relatively they will wait for \$110 as opposed to receiving the \$100 today.
+This increased focus on how long they will wait, leads to decreased patience when times are framed as ratios.
 
-This work aims to demonstrate that the fourth research stream -- that patience is affected by how options are framed -- can affect real world phenomenon.
-Specifically, this paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience.
+I test the ETRE across three studies. 
+The Study 1 tests the ETRE in standard intertemporal choices framed as gains.
+As a test of robustness, Study 2 tests how the ETRE effects patience in losses. 
+Although losses have been shown to be a critical moderator of intertemporal preferences, few studies explore how losses affect intertemporal choices \cite{Hardisty2013, Hardisty2015}.
+Further all studies cited so far have only looked at framing effects of intertemporal choices framed as gains \cite{Magen2008, Read2005, Read2013}. 
+Framing effects which are robust for gains, may affect losses differently.
 
-Specifically, I focus on how framing times as ratios can decrease patience.
-When delays are sufficiently small, and when temporal ratios are made explicit, our attention will be focused on the ratio of how much \textit{longer} we will wait. 
-For instance reframing a choice between ``receiving \$100  in 1 day'' and ``\$110 in 10 days'' to being a choice between ``receiving \$100  in 1 day'' and ``\$110 in 10 \textit{times as long}'', will focus participants on how much longer they will wait for \$110 as opposed to receiving the \$100 today.
-This increased focus on how long they will wait, leads to decreased patience when times are framed as ratios.
+Study 3 expands on Studies 1 and 2 by testing the ETRE in an explicit shipping scenario. 
+Shipping scenarios are formatted differently than intertemporal choices in the previous studies.
+Study 3 serves as test that the ETRE happens in explicit shipping scenarios.
+
+In what follows I explicitly state three hypotheses generated to test the ETRE. 
+I then present three studies to test these hypotheses. 
+Finally, I conclude with the discussion and implications of the ETRE.
 
@@
 \subsubsection{Hypothesis 1}
+Gain intertemporal choices take the form of receiving differing amounts of money at two times, e.g.  ``receiving \$100  in 1 day'' and ``receiving \$110 in 10 days''.
 Given that people rely on differences and ratios between amounts and times in intertemporal choice, increasing the reliance on ratios for times with relatively small differences should lead to decreased patience.
@@
 \subsubsection{Hypothesis 2}
+
+Loss intertemporal choices take the form of paying differing amounts of  money at two times, e.g.  ``paying \$100  in 1 day'' and ``paying \$110 in 10 days''.
 Similar to hypothesis 1 when framed as losses making temporal ratios explicit people will be more impatient.
@@
 \subsubsection{Hypothesis 3}
-Related to hypothesis 2 when framed an intertemporal chioce is framed as a shipping scenario,  making temporal ratios explicit people will induce less patience.
+Although choosing a shipping speed is similar to a loss framed intertemporal choice, there is a subtle difference between the two.
+When selecting a shipping speed people are asked to pay now to receive something sooner, whereas for the items in Study 2 people are asked to pay at two different times. 
+However there is no good reason to suspect that the ETRE would change in the shipping scenario.
+Related to hypothesis 2 when framed an intertemporal choice is framed as a shipping scenario,  making temporal ratios explicit people will induce less patience.
 
@@
 
-In this paper, I describe three studies investigating the explicit temporal ratio effect (ETRE).
-Study 1 shows that the explicit temporal ratios lead to less patience when intertemporal choices are framed as gains.
-Study 2 extends this effect to losses.
+Study 1 shows that the explicit temporal ratios lead to less patience in intertemporal gains.
+Study 2 extends this effect to intertemporal losses.
 Finally, Study 3 shows that in an explicit shipping scenario people are less patient when temporal ratios are made explicit.
 
-%Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
-%Study 2 investigates how framing times as ratios leads to decreased patience in losses. 
-%
-%Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
-%Shipping scenarios are typically displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
-%Although choosing a shipping speed is similar to the previous intertemporal choices there is a subtle, yet possibly significant difference between the two.
-%Shipping decisions can be thought of as \textit{paying more now to receive an item sooner} whereas the intertemporal choices shown in Study 2 can be thought of as \textit{paying something smaller now to avoid paying something larger later}.
-%This difference in choices  
-%
-
 \section{Study 1}
-Study 1 shows that the explicit temporal ratios lead to less patience when intertemporal choices are framed as gains.
+Study 1 makes implicit temporal ratios explict in intertemporal gains.
 Participants chose between smaller sooner and larger later options. 
@@
 \subsection{Method}
-Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
+Two hundred participants with standard demographics from Amazon Mechanical Turk were recruited to take part in a ``Decision Making Survey''. 
 Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
@@
 
-For the implicit time ratio condition, intertemporal choice A was displayed as follows:
+For participants in the implicit time ratio condition, intertemporal choice A was displayed as follows:
 
@@
 
-In the explicit time ratio condition, intertemporal choice A was displayed as follows:
+For participants in the explicit time ratio condition, intertemporal choice A was displayed as follows:
 
@@
 
-Further each choice was displayed in a random order to each participant. 
 After completing their choices participants completed demographic information.
@@
 
-Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis tests.
- Overall the effect of ratio was significant. 
+Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis test.
+Explicit ratio participants made significantly more sooner choices than implicit ratio participants. 
 To confirm these results, I ran a robust logistic regression which allowed for correlated errors within participants.
@@
 Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
+Participants who participated in Study 1 were not allowed to participate in Study 2. 
 Participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
@@
 Participants completed informed consent, then were given a test to ensure they understood the paradigm.
+Next, participants read the instruction
+``Please imagine you face a set of choices
+about paying a bill at a short time from now, or another bill later''.
 Next they were assigned to one of the two conditions: implicit ratio or explicit ratio.
-Participants read the instruction
-``Please imagine you face a set of choices
-about paying a bill at a short time from now, or another bill later''
+
 They then were shown the five intertemporal choices in from the Study 2 section of  Table~\ref{tab:stimuli}. 
@@
 
-Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis tests.
+Statistical significance of  the difference between explicit and implicit temporal ratios was tested via a Kruskal-Wallis test.
 Overall the effect of ratio was significant. 
@@
 Although choosing a shipping speed is similar to the intertemporal choices shown in Studies 1 and 2 there is a subtle, yet possibly significant difference between the two.
-Shipping decisions can be thought of as \textit{paying more now to receive an item sooner} whereas the intertemporal choices shown in Study 2 can be thought of as \textit{paying something smaller now to avoid paying something larger later}.
+Shipping decisions can be thought of as \textit{paying more now to receive an item sooner} whereas the intertemporal choices shown in Study 2 can be thought of as \textit{paying something smaller sooner to avoid paying something larger later}.
 
@@
 Participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study. 
-Additionally only participants who passed an attention check were included in the final analysis.
+Further, participants who had participated in either Study 1 or Study 2 were not allowed to take Study 3. 
+Additionally, only participants who passed an attention check were included in the final analysis.
 These precautions were taken to ensure that the sample included qualtiy participants who were paying attention to the study.
@@
 You are about to check out and are presented with two shipping options.''
-They then were shown the five intertemporal choices in from the Study 3 section of  Table~\ref{tab:stimuli}. 
+They then were shown the five shipping choices in from the Study 3 section of  Table~\ref{tab:stimuli}. 
 Choices were displayed in a randomized order
@@
 As with Studies 1 and 2, the dependent variable was the number of times the participant chose sooner reward (option 1 in Table~\ref{tab:stimuli}). 
-As seen the Table~\ref{tab:study3results} (note I did not actually collect data for this study I just put sample data down for illustrative purposes) there is a strong effect of explicit intertemporal ratios.
+As seen the Table~\ref{tab:study3results} (NOTE: I did not actually collect data for this study I just put sample data down for illustrative purposes) there is a strong effect of explicit intertemporal ratios.
 Nearly twice as many people choose to expedite their shipping temporal ratios are explicit.
@@
 \begin{table}[!ht]
-	\caption{Proposed Percent of choices in Losses for the sooner option in different temporal ratios} 
+	\caption{Proposed Percent of choices in shipping for the sooner option in different temporal ratios} 
 	\label{tab:study3results}
@@
 		\hline
-		A (\%)	 & 20  & 48\\
-		B (\%) 	 & 25  & 50\\
-		C (\%) 	 & 27  & 57\\
-		D (\%)   & 28  & 54\\
-		E (\%)   & 24  & 58\\
+		A (\%)	 & 30  & 58\\
+		B (\%) 	 & 40  & 60\\
+		C (\%) 	 & 31  & 61\\
+		D (\%)   & 42  & 55\\
+		E (\%)   & 38  & 68\\
 		Mean (\%) & 38  & 65.8 \\
@@
 
-Across three studies we demontrate that the Explicit Temporal Ratio Effect (ETRE) -- making temporal ratios explicit -- yields greater patience. 
-This importance has been understated by previous models of intertemporal choice, and by explicitly expressing temporal ratios people are more impatient.
+Across three studies we demonstrate that the Explicit Temporal Ratio Effect (ETRE) -- making temporal ratios explicit -- yields greater patience. 
+This result holds for gain and loss intertemporal choices as well as an explicit shipping scenario.
 
 \section{Limitations and Future Directions}
+
+All of the intertemporal choices made in this paper were hypothetical.
+However studies have found that people make identical choices in hypothetical and real outcomes \cite{Johnson2002, Bickel2009}.
+
+Further all intertemporal choices were made with on a relatively short timeframe.
+The ETRE could reverse in the following choice ``Receive \$10 in 100 days'' and ``Receive \$15 in 120 days'' is reframed as  ``Receive \$10 in 100 days'' and ``Receive \$15 in 1.2 times as long''.
+
+Related to the previous point all ratios presented in this study were whole integers.
+Since the real world is rarely so neat as the laboratory, future work investigating how people respond to decimals in explicit ratios is an important next step for this research.
 

### 2015-12-18 wave_10dec/presentation/methods_results_18dec.Rmd
Analyze data and methods section

---
title: "Ratio Differences in Unlikeness and Choice"
author: "Daniel Wall"
date: "December 18, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r, echo=FALSE, message = FALSE}
setwd("/git_repositories/timeratios/wave_10dec/")
library(knitr)
library(xtable)
source("load_packages.R")
rm(list = ls())
load("analyzed_data/analyzed_data.Rdata")

opts_knit$set(echo=FALSE, message = FALSE, warning = FALSE)
```

# Study 1: Does a large change in temporal ratios alter similarity and choice.

## Method

As a requirement for course credit, `r nrow(timerat)` Rutgers undergradutates took part in this study. 
This study was taken along with six other, unrelated, studies which will not be reported furter. 
After participants completed informed consent, they were assigned to one of two temporal ratio change conditions: 1) Small temporal ratio change and 2) Large temporal ratio change. 
Stimuli can be seen in Table [].

![alt text](images/stimulitable.png)

Participants then made unlikeness judgements between two times on a slider bar with a range of 0 - 100. 
Specifcally they were told: "For the first part of this study we will ask you how "unlike" you perceive two different points in time to be.
You will be given a slider bar with:
Not at all Unlike each other on the left and 
Completely Unlike each other on on the right"

Next within the ratio change condition the the delayed or not delayed block was shown first followed by the other block. 
This was done to minimize the chance that related choices were shown sequentially. 
Within each block choices were presented in a random order.

After completing 12 unlikeness judgments participants completed an unrelated task which will not be reported further. 
Next participants completed 12 intertemporal choice questions.

Similar to the unlikeness judgments participants either saw the delayed or not delayed block of choices first. 
Once again within a block choices were randomly presented to subjects. 
The monetary amounts for the choices can be seen in the last two rows of Table [].

## Results

### Unlikeness Judgments

First, for each choice we subtracted the "not delayed"" unlikeness judgment from the "delayed" unlikeness judgment to calculate the unlikeness difference measure. 
Figure [] shows the unlikeness difference measure for each ITC id for both the large and small ratio temporal ratio change conditions along with 95% confidence intervals.

```{r, echo=FALSE, message = FALSE}
ggplot(diss1_sum, aes(x = id_pair, y = mean_dissdiff, fill = ratcng, color = ratcng)) +
  geom_bar(stat="identity", position = "dodge") +
  geom_errorbar(aes(ymin = lower, ymax = upper), position = "dodge")

```

We also ran non-paired t tests to check for significant differences which can be seen in Table[].
There are no significant differences between the large and small ratio changes for any id pair.

```{r, echo=FALSE, message = FALSE}
diss_t_tests
```

### ITC Choices

Below are the percentage of larger later choices for each of the ITC ids in both the delayed and not delayed conditions.
The top panel is the large temporal ratio change condition and the bottom panel is the small temporal ratio change condition.

```{r, echo=FALSE, message = FALSE}
ggplot(itc_sum_id, aes(x = id_pair, y = estimate, group = delayed, fill = delayed)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = conf.high, ymin = conf.low), position = "dodge") +
  facet_grid(ratcng ~ .)
```

Next, for each intertemporal choice id, we ran a logistic regression predicting choice with 0 = Smaller Sooner and 1 = Larger Later. 
The indepent variables were whether or not the choice was delayed, the change in temporal ratio, and the interaction between the two. 
The figure below shows the regression coefficients along with the 95% CI.

```{r, echo=FALSE, message = FALSE}
ggplot(itc_glms_sub, aes(x = id_pair, y = estimate, group = term,  fill = term)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = upper, ymin = lower), position = "dodge")

```

As expected, regardless of delay, people are significantly more likely to choose the larger later option when the ratio change is small.
There is no effect of whether or not the the choice was delayed. 
Futher there was no interaction between delayed and the type of ratio change.

These results were confirmed by an omnibus logistic regression predicting choice with delay type, ratio change type, each unique intertemporal choice, and all possible interactions. 
As seen by the ANOVA below.
There is a main effect of temporal ratio change and of each unique intertemporal choice, but no other effects approach significance.

```{r, echo=FALSE, message = FALSE}
Anova(all_int_glm, type = 3)

```

### 2015-12-18 wave_10dec/0_gather_clean_data.R

@@ -79,14 +79,48 @@ ggplot(eff_df, aes(x = ratcng, y = fit)) +
   geom_errorbar(aes(ymin = lower, ymax = upper))
 
 
-# now lets look at the intertemporal choices
+#### now let's look at pairs of choices
+
+dissc1 <- dissc %>%
+  mutate(id_pair = ifelse(id > 6, id - 6, id))
+
+
+# plot it
+diss1_sum <- dissc1 %>%
+  group_by(ratcng, id_pair) %>%
+  summarise(mean_dissdiff = mean(diss_diff, na.rm=TRUE),
+            se_dissdiff = std.error(diss_diff, na.rm=TRUE)) %>%
+  mutate(lower = mean_dissdiff - 1.96*se_dissdiff,
+         upper = mean_dissdiff + 1.96*se_dissdiff)
+
+ggplot(diss1_sum, aes(x = id_pair, y = mean_dissdiff, fill = ratcng, color = ratcng)) +
+  geom_bar(stat="identity", position = "dodge") +
+  geom_errorbar(aes(ymin = lower, ymax = upper), position = "dodge")
+
+
+# do t tests
+diss_t_tests <- dissc1 %>%
+  group_by(id_pair) %>%
+  do(tidy(t.test(diss_diff ~ ratcng, data = .)))
+
+
+
+##### now lets look at the intertemporal choices #####
 itc <- timerat[, str_detect(names(timerat), "^(responseid|itc\\d+_\\d+)")] %>%
   melt(id = "responseid") %>%
   filter(!(is.na(value))) %>%
   mutate(ss_time = str_extract(variable, "\\d+"),
          ll_time = str_extract(variable, "\\d+$")) %>%
   group_by(responseid) %>%
-    merge(diss_id, by = c("ss_time", "ll_time"))
+    merge(diss_id, by = c("ss_time", "ll_time")) %>%
+  mutate(value = factor(value),
+         diss_set = factor(diss_set),
+         delayed = factor(delayed),
+         id_pair = ifelse(id > 6, id - 6, id),
+         ratcng = ifelse(diss_set == 1, "small",
+                         ifelse(diss_set == 2, "large", NA)) %>%
+           factor)
+
   
 #summarise and plot the data
 itc_sum <- itc %>%
@@ -110,10 +144,14 @@ ggplot(itc_sum,
   
 
 # now lets run a model on this
-itc <- itc %>%
-  mutate(value = factor(value),
-         diss_set = factor(diss_set),
[…]
